---
title: ""
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

```{=html}
<style>

body {
    font-family: "Times New Roman";
    font-size: 11.5pt;
    color: black;
}

.title {
  font-family: "Times New Roman";
  font-size: 16pt;
  font-weight: bold;
  color: black;
}

h1 {
    font-family: "Arial";
    font-size: 16pt;
    font-weight: bold;
    color: black;
}


h2 {
  font-family: "Arial";
  font-size: 11.5pt;
  font-weight: bold;
  color: black;
}

h3 {
  font-family: "Arial";
  font-size: 10.5pt;
  font-weight: bold;
  color: black;
}

h4 {
    font-family: "Times New Roman";
    font-size: 12pt;
    color: black;
}

p {
    font-family: "Times New Roman";
    font-size: 11.5pt;
    color: black;
}

.caption {
    font-family: "Arial";
    font-size: 10pt;
    color: black;
}

th {
    font-family: "Arial";
    font-size: 9pt;
    font-weight: bold;
    color: black;
}

tr,td,table {
    font-family: "Arial";
    font-size: 9pt;
    color: black;
}

.header{
  font-family: "Arial";
  font-size: 10pt;
  font-weight: bold;
  color: black;
}

</style>
```

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) # Prevent overwhelming messages from reaching user

options(download.file.method = "wininet")
#install.packages("sqldf")
# Load libraries
library(RODBC)
library(knitr)
library(ggplot2)
library(sp)
library(sf)
library(sqldf)
library(tidyverse)


# Make kable show blanks for NAs
options(knitr.kable.NA = '')

# These parameters restrict the dimensions of plots to 6 metric inches so they don't blow out Word
FigWidth = 6
```

```{r echo=FALSE,label="SET UP REPORT PARAMETERS HERE"}

# Set up report parameters here
SurveyYear = 2024  # Set this to the desired report year. Don't set to 2016 because that survey didn't come off and the script will bomb

# ODBC connection to the BAEA monitoring SQL Server Database. 
# Contact data manager for server and database details.
# ConnectionString to the bald eagles Sql Server database
BaldEaglesConnectionString = "Driver=SQL Server Native Client 11.0;Server=inpyugamsvm01\\nuna;Database=CAKN_BaldEagles;trusted_connection=YES"

# Database connection to the eagles database
BaldEaglesConnection = odbcDriverConnect(connection = BaldEaglesConnectionString)

# Set the path to shapefiles used in the map, AK, Parks, Rivers, etc
ShpPath = "C:/Work/GIS Common Layers/"

```



```{r echo=FALSE,label="Functions"}

# Current date
CurrentDate = format(Sys.Date(), format='%B %d, %Y') # Get the current date

# Function to return data frames from SQL database queries
GetDataFrameFromDatabase <- function(Sql) {
   return(sqlQuery(BaldEaglesConnection, Sql))
}

# Function to convert a date string to a long date
GetLongDate <- function(Date) {
  return(format(as.Date(Date), format="%B %d, %Y"))
}

# These counters are used to number the tables and figures. Each gets incremented after use.
TableCounter = 1
FigureCounter = 1

```


# Post-survey Data Quality Report For The Bald Eagles Monitoring Program in Wrangell-St. Elias National Park and Preserve, Alaska

Scott D. Miller  
National Park Service, Arctic Inventory and Monitoring Network  
240 W. 5th Ave  
Anchorage, Alaska 99501   

`r CurrentDate`

U.S. Department of the Interior\
National Park Service, Alaska Region\
Anchorage, Alaska\

# Introduction

In the realm of biological sciences, the integrity of data is crucial for advancing our understanding of complex biological processes and ecosystems. Ensuring high data quality is fundamental for producing reliable results, validating hypotheses, and making informed decisions for the management of natural areas, enhancing conservation efforts, and communicating results to the public. Data quality control is a critical step in the scientific life cycle for long-term monitoring of ecosystems, ensuring the accuracy, consistency, and completeness of data collected. This report documents the quality control procedures executed on the field data collected during the `r SurveyYear` bald eagles monitoring survey in Wrangell-St. Elias National Park and Preserve in Alaska.

Each section of this report details a quality control check performed on the master bald eagles monitoring database, along with any findings. Any identified quality control defects should be fixed or documented. The report is broken out into two major sections related to 1) information about eagle nests and 2) survey results.

# Quality Control Tests: Nests

## QC-1: Active nests with DateDiscoveredGone is not NULL

All active nests should have DateDiscoveredGone attributes NULL

```{r echo=FALSE}
Sql = paste("SELECT Nest_ID, Date_O, Nest_O, Date_P, StatusCode_P, Comments_O, Comments_P, IsActive, DateDiscoveredGone
FROM QC_ActiveNestsWithDateDiscoveredGoneIsNotNULL
WHERE Year(Date_O) = ",SurveyYear,sep='')
df = GetDataFrameFromDatabase(Sql)
kable(df)

```


`r Sql` 

`r nrow(df)` rows

## QC-2: Nests With No Productivity Observation

Occasionally observers are unable to re-survey nests for the productivity portion of the survey due to bad weather, smoke, logistical, or other problems. These records should have their StatusCode_P attribute set to 'NS' (Not surveyed). The table below shows records with StatusCode_P = 'NS'

```{r echo=FALSE}
Sql = paste("SELECT Nest_ID, RiverSegment, Date_O, OccupancyStatus, ProductivityStatus,Comments_O,Comments_p
FROM QC_NestsNotProductivitySurveyed
WHERE (NOT (Date_O IS NULL)) AND (Date_P IS NULL)
And Year(Date_O)=",SurveyYear," and StatusCode_P='NS'
ORDER BY Date_O,Nest_ID",sep='')
df = GetDataFrameFromDatabase(Sql)
kable(df)

```

`r Sql`

`r nrow(df)` rows



<!-- ## QC: Inactive nests that possibly should be active -->

<!-- The following nests, if any, should possibly be inactivated. -->

<!-- ```{r echo=FALSE} -->
<!-- Sql = paste("SELECT  [Nest_ID] -->
<!--       ,[IsActive] -->
<!--       ,[NumberOfTimesNotFoundInTheLast2Surveys] -->
<!--       ,[LastSightingDate] -->
<!--       ,[Comments] -->
<!-- --      ,[SELECTQuery] -->
<!-- --     ,[NestsQuery] -->
<!-- --      ,[UPDATEQuery] -->
<!--   FROM [CAKN_BaldEagles].[dbo].[QC_Nests_InActiveNestsThatPossiblyShouldBeActive] -->
<!-- ORDER BY LastSightingDate",sep='') -->
<!-- df = GetDataFrameFromDatabase(Sql) -->
<!-- kable(df) -->

<!-- ``` -->

<!-- `r Sql` -->

<!-- `r nrow(df)` rows -->

## QC-3: Active nests that possibly should be inactive

The following nests, if any, should possibly be inactivated.

```{r echo=FALSE}
Sql = paste("SELECT        Nest_ID, IsActive, DiscoverDate, LastSightingDate, NumberOfTimesNotFoundInTheLast2Surveys, Comments
--, HelperQuery, UpdateQuery
FROM            QC_NestsThatProbablyShouldBeInactivated",sep='')
df = GetDataFrameFromDatabase(Sql)
kable(df)

```

`r Sql`

`r nrow(df)` rows







# Quality Control Tests: Survey Observations

## QC-4: Active nests that were not search for during the latest survey

The results below active nests, if any, that were not searched for during the latest survey.

```{r echo=FALSE}
Sql = paste("SELECT [Nest_ID]
      ,[RiverSegment]
      ,[LastSightingDate]
      ,[IsActive]
      ,[NumberOfTimesNotFoundInTheLast2Surveys]
      ,[Comments]
 --     ,[UPDATEQuery]
--      ,[HelperQuery]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_ActiveNestsNotSearchedForDuringLastSurvey]",sep='')
df = GetDataFrameFromDatabase(Sql)
kable(df)

```

`r Sql`

`r nrow(df)` rows


## QC-5: Records from the latest survey that have initiated nests, but no follow up productivity information

The following records, if any, indicate that nests were observed as initiated, but no productivity information exists.

```{r echo=FALSE}
Sql = paste("SELECT [Year]
      ,[Total records]
      ,[Number of initiated nests]
      ,[Initiated nests with no follow up survey]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_MissingLateProductivitySurvey_Summary] 
  WHERE Year=",SurveyYear,sep='')
df = GetDataFrameFromDatabase(Sql)
kable(df)

```

`r Sql`

`r nrow(df)` rows



## QC-6: Possibly mistaken nest initiation decisions

Initiated nests should have at least two adults present during the occupancy survey with incubating behaviors observed. The following observations may have been mistakenly ruled as initiated nests since these rules were violated and the nests did not result in young.

```{r echo=FALSE}
Sql = paste("SELECT [Year]
      ,[Nest_ID]
      ,[Date_O]
      ,[Nest_O]
      ,[NumEggs_O]
      ,[NumAds_O]
      ,[AdBehav1_O]
      ,[AdBehav2_O]
      ,[NumYng]
      ,[StatusCode_P]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_PossiblyBadNestInitiations]
  WHERE Year=",SurveyYear," ORDER BY Nest_ID",sep='')
df = GetDataFrameFromDatabase(Sql)
kable(df)

```

`r Sql`

`r nrow(df)` rows


## QC-7: Nests marked successful that did not produce any young

Nests marked successful that did not produce any young.

```{r echo=FALSE}
Sql = paste("SELECT [Nest_ID]
      ,[NumYng]
      ,[StatusCode_P]
      --,Year
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_SuccessfulNestsWithNoYoung] WHERE Year=",SurveyYear," ORDER BY Nest_ID",sep='')
df = GetDataFrameFromDatabase(Sql)
kable(df)

```

`r Sql`

`r nrow(df)` rows



## QC-8: Chick Age Recorded But No Young Recorded

The following records, if any, have a Chick_Age attribute that is not NULL, but NumYng attribute is NULL or zero. This means a chick stage was recorded, but no chicks were observed.

```{r echo=FALSE}
Sql = paste("SELECT        Year, Nest_ID, Date_O, Nest_O, Date_P, StatusCode_P, NumYng, Chick_Age, Comments_O, Comments_P
FROM            QC_ChickAgeButNoYoung WHERE Year = ",SurveyYear,sep='')
df = GetDataFrameFromDatabase(Sql)
kable(df)

```

`r Sql`

`r nrow(df)` rows


## QC-9: Successful Nests With No Recorded Young

The definition of nest success is observed young during the productivity survey. The following records, if any, should be checked.

```{r echo=FALSE}
Sql = paste("SELECT [Year]
      ,[Nest_ID]
      ,[NumYng]
      ,[StatusCode_P]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_SuccessfulNestsWithNoYoung]
  WHERE Year = ",SurveyYear,sep='')
df = GetDataFrameFromDatabase(Sql)
kable(df)

```

`r Sql`

`r nrow(df)` rows



## QC: List of comments

Comments can be useful in troubleshooting strange results

```{r echo=FALSE}

df = GetDataFrameFromDatabase(paste("SELECT Nest_ID,Comments_O as Comment,'Occupancy' as Survey  
FROM SurveyData 
WHERE Year = ",SurveyYear," And (NOT (Comments_O IS NULL))
UNION
SELECT Nest_ID,Comments_P,'Productivity' 
FROM SurveyData 
WHERE Year = ",SurveyYear," And (NOT (Comments_P IS NULL))
ORDER BY Nest_ID",sep=""))

kable(df)

```

## R Markdown Report Template

This template was generated automatically from an R Markdown script available at <https://github.com/NPS-ARCN-CAKN/WRST-BAEA-Annual-Report-R-Markdown>. The information in the report was generated from bald eagle survey data in the NPS WRST_Caribou SQL Server database (inpyugamsvm01\nuna:WRST_Caribou).

