---
title: ""
format:
  typst:
    toc: false
    number-sections: false
    columns: 1
  html: default
---

```{r setup, include=FALSE, echo=FALSE}

# Set up report parameters here
SurveyYear = 2025  # Set this to the desired report year. Don't set to 2016 because that survey didn't come off and the script will bomb

knitr::opts_chunk$set(warning = FALSE, message = FALSE) # Prevent overwhelming messages from reaching user
options(download.file.method = "wininet")

# Make kable show blanks for NAs
options(knitr.kable.NA = '')

# Load libraries
library(odbc)
library(knitr)
library(sf)
library(tidyverse)

# Connection to the bald eagles Sql Server database
Connection = dbConnect(odbc(),Driver = "Sql Server",Server = "INP2300IRMADB01.nps.doi.net\\NTWK", Database = "CAKN_BaldEagles")

# Current date
CurrentDate = format(Sys.Date(), format='%B %d, %Y') # Get the current date

# Function to convert a date string to a long date
GetLongDate <- function(Date) {
  return(format(as.Date(Date), format="%B %d, %Y"))
}

# These counters are used to number the tables and figures. Each gets incremented after use.
TableCounter = 1
FigureCounter = 1

# Load spatial layers from either Data Store (slow, but reliable for anyone anywhere) or local directory
if(1==1){
  # Load spatial files locally
  ShpPath = "C:/Work/GIS Common Layers/"
  Parks = read_sf(paste(ShpPath,"GCS/AKParks_reprojected.shp",sep=""))
  AK  = read_sf(paste(ShpPath,"GCS/AK_GCS_poly.shp",sep=""))
  Rivers = read_sf(paste0(ShpPath,"Alaska_Rivers_1M.geojson.geojson"))
  Roads = read_sf(paste0(ShpPath,"AK_Roads.geojson"))
  Lakes = read_sf(paste0(ShpPath,"AK_Lakes/mv_major_lake_py.shp"))
}else{
  # Load spatial files from Data Store
  Parks = st_read('https://irma.nps.gov/DataStore/DownloadFile/704836')
  AK = st_read('https://irma.nps.gov/DataStore/DownloadFile/746052')
  Rivers = read_sf('https://irma.nps.gov/DataStore/DownloadFile/758201') 
}
#ggplot() + geom_sf(data=Lakes)

```

```{r echo=FALSE, label="Survey dates"}

# Get a summary of survey results, omitting 2016 which was a bust survey
Sql = paste("SELECT [Year]
,[Records]
,[Available nests]
,[Available but omitted]
,[Nests located]
,[Nests not located]
,[Nests not counted]
,[Productivity unknown]
,[Productivity known]
,[New]
,[Lost]
,[Net gain]
,[Ending nest count]
,[Occupied]
,[Pct. occupied]
,[Successful]
,[Pct. Successful]
,[Eggs]
,[Eggs per occupied nest]
,[Young]
,[NestsProducingOneYoung]
,[NestsProducingTwoYoung]
,[NestsProducingThreeYoung]
,[NestsProducingGTThreeYoung]
,[Young per successful nest]
,[Fledglings]
,[Fledglings per successful nest]
,[Start of occupancy surveys]
,[End of occupancy surveys]
,[Number of occupancy survey days]
,[Start of productivity surveys]
,[End of productivity surveys]
,[Number of productivity survey days]
,[Certified records] FROM Dashboard WHERE Year <= ",SurveyYear," AND Year NOT IN (2016) 
ORDER BY [Year] DESC",sep="")
Summary = dbGetQuery(Connection,Sql) 

# Build some survey date values to use later in the markdown text
OStart = Summary %>% filter(Year == SurveyYear) %>% pull(`Start of occupancy surveys`) 
OEnd = Summary %>% filter(Year == SurveyYear) %>% pull(`End of occupancy surveys`) 
PStart = Summary %>% filter(Year == SurveyYear) %>% pull(`Start of productivity surveys`) 
PEnd = Summary %>% filter(Year == SurveyYear) %>% pull(`End of productivity surveys`) 

# Length of record
MinYear = min(Summary$Year, na.rm = TRUE) 
MaxYear = max(Summary$Year, na.rm = TRUE)
LengthOfRecord = MaxYear - MinYear
FirstSurveyYear = min(Summary$Year)
NumberOfSurveys = Summary %>% summarize(n=n()) %>% pull(n) 

# Build some summary statistics to report in the markdown
NestsLocated = Summary %>% filter(Year == SurveyYear) %>% pull(`Nests located`) 
ProductivityKnown = Summary %>% filter(Year == SurveyYear) %>% pull(`Productivity known`) 
Occupied = Summary %>% filter(Year == SurveyYear) %>% pull(Occupied) 
Successful = Summary %>% filter(Year == SurveyYear) %>% pull(Successful) 
YoungFledged = Summary %>% filter(Year == SurveyYear) %>% pull(Young) 
NF = Summary %>% filter(Year == SurveyYear) %>% pull(`Nests not located`) 
NC = Summary %>% filter(Year == SurveyYear) %>% pull(`Nests not counted`) 
NewNests = Summary %>% filter(Year == SurveyYear) %>% pull(New) 
LostNests = Summary %>% filter(Year == SurveyYear) %>% pull(Lost) 
YngPerSuccessfulNest = signif(Summary %>% filter(Year == SurveyYear) %>% pull(`Young per successful nest`) ,3)
PctOccupied = signif(Summary %>% filter(Year == SurveyYear) %>% pull(`Pct. occupied`) ,2)
PctSuccessful = signif(Summary %>% filter(Year == SurveyYear) %>% pull(`Pct. Successful`) ,2)
ProductivityUnknown = Summary %>% filter(Year == SurveyYear) %>% pull(`Productivity unknown`)

ExecutiveSummaryText = paste("Nest occupancy surveys were flown between ",GetLongDate(OStart)," and ",GetLongDate(OEnd),", followed by productivity surveys between ",GetLongDate(PStart)," and ",GetLongDate(PEnd),". We observed ",NestsLocated," nests, of which ",Occupied," were occupied (",PctOccupied,"%). Productivity surveys revealed that ",Successful," occupied nests were successful (i.e., produced young; ",PctSuccessful,"%). Successful nests produced ",YoungFledged," young, an average of ",YngPerSuccessfulNest," young per successful nest.",ifelse(NewNests == 1," One new nest was ",paste0(" ",str_to_sentence(xfun::numbers_to_words(NewNests))," new nests were"))," discovered." ,sep="")

```

# Survey Summary: Bald Eagle Monitoring at Wrangell-St. Elias National Park and Preserve, `r SurveyYear`

Kyle Cutting^1^ and Scott Miller^2^

^1^ National Park Service   
Wrangell St. Elias National Park and Preserve   
8 Richardson Hwy Mile 106   
Copper Center, AK 99573

^2^ U.S. Department of the Interior   
240 W. 5th Ave Anchorage   
Alaska 99501   

`r CurrentDate`

# Abstract

The Copper River drainage in Wrangell-St. Elias National Park and Preserve (WRST), Alaska, hosts a high density of nesting Bald Eagles *(Haliaeetus leucocephalus)*. Bald Eagles in WRST are a high profile species that are dependent upon many resources, including salmon, along the Copper River and nest at the northern edge of the their range. Bald Eagles are top-trophic level predators and they often respond quickly to changes in their environment by changing their breeding activities.

In `r SurveyYear` we conducted an aerial bald eagle productivity survey in and around WRST as part of long-term monitoring. `r ExecutiveSummaryText`

# Objectives

Document the annual changes in occupancy, productivity, and spatial distribution of bald eagles,

# Methods

The bald eagle survey is part of long-term monitoring since 2004 (Putera and Miller, 2018). Five areas covering 378 miles of riverine habitat along the Copper, Chitina, and Bremner Rivers were surveyed (Figure `r FigureCounter`).

Two aerial surveys following the river corridor were conducted in 2025 - one to document territorial occupancy, and the other to document productivity. Surveys were flown using a small airplane flying at low level (<100m AGL) and low speed (<100 kph), with one experienced observer. 

We collected the following data on each nest observed: number of adults, behavior of adults, number of eggs and/or nestlings, tree species, status, and nest location. All nests still intact from the previous year’s surveys were rechecked for occupancy. During the productivity (chick) survey, all previously occupied nests from the first survey were rechecked . We define occupancy as a bird in incubation posture or two adult birds at a nest or in the nest tree. For the productivity survey, we rechecked all occupied nests from the first survey to see if chicks were present.


```{r echo=FALSE}

# Build a map of the survey area showing park boundaries, copper river and tributaries, etc.
ggplot() +
  
  # Blue rectangle to make the background of the map blue
  geom_rect(aes(xmin = -155,xmax = -140,ymin = 55,ymax = 70),fill = "steelblue3") +
  
  # Contextual layers
  geom_sf(data=AK,mapping=aes(), color='darkgray',fill='snow2') +
  geom_sf(data=Parks,mapping=aes(), color='yellow4',fill='yellow4',alpha=0.3) +
  geom_sf(data=Rivers %>% filter(TYPE != 'I'),mapping=aes(),color='steelblue3',fill='steelblue3') +
  geom_sf(data = Roads, fill = "gray50", color = 'gray50',alpha=0.4) +
  geom_sf(data=Lakes,mapping=aes(),color='steelblue3',fill='steelblue3') +
  
  # Set the map limits, zoom into the BAEA monitoring area
  coord_sf(xlim = c(-146.6, -142), ylim = c(60.5, 62.8), expand = FALSE) +
  xlab('Longitude') +
  ylab('Latitude') +
  

  
  # WRST text Label
  geom_text(aes(x = -144.25, y = 61.8, label = "Wrangell - St. Elias National\nPark and Preserve"), size = 4, color = "gray40", fontface = "bold") +
  
  # Upper Copper River label
  geom_text(aes(x = -144, y = 62.45, label = "Upper Copper River"), size = 4, color = "black") +
  
  # Middle Copper River label
  geom_text(aes(x = -145.2, y = 61.5, label = "Middle Copper River"), size = 4, color = "black") +
  
  # Lower Copper River label
  geom_text(aes(x = -145.2, y = 60.8, label = "Lower Copper River"), size = 4, color = "black") +
  
  # Chitina River label
  geom_text(aes(x = -143.8, y = 61.25, label = "Chitina River"), size = 4, color = "black") +
  
  # Bremner River label
  geom_text(aes(x = -143.8, y = 60.95, label = "Bremner River"), size = 4, color = "black") +
  
  # Minimal theme
  theme_minimal() +
  
  # Legend position
  theme(legend.position = "bottom")


```
Figure `r FigureCounter`. Bald eagle monitoring study area.

# Results

Table `r TableCounter` shows overall nest success and productivity patterns observed in `r SurveyYear`. Table `r TableCounter + 1` lists productivity metrics by river segment. Spatial distribution of surveyed nests and their productivity is shown in Figure `r FigureCounter + 1`.

Table `r TableCounter`. Results summary, `r SurveyYear` bald eagle survey.

```{r echo=FALSE, label="Current survey metrics table"}
# Increment the figure counter
FigureCounter = FigureCounter +1

# Summarize the eagle counts for the current survey into 
# a table, transpose it so it fits the report better, modify the header, and output
Summary_Current = Summary %>% 
  filter(Year == SurveyYear) %>% 
  mutate(
    `Pct. occupied` = round(`Pct. occupied`,1),
    `Pct. Successful` = round(`Pct. Successful`,1),
    `Young per successful nest` = round(`Young per successful nest`,2)
  ) %>%
  select(    
    `Nests located`,
    `Occupied`,
    `Successful`,
    `Productivity unknown`,
    `New`,
    `Lost`,
    
    `Pct. occupied`,
    `Pct. Successful`,
    `Young`,
    `Young per successful nest`,
    -`Net gain`,
    -`Available nests`,
    -`Available but omitted`,
    -`Nests not located`,
    -`Nests not counted`,
    -`Ending nest count`,
    -Year,
    -Records,
    -`Certified records`,
    -`Start of occupancy surveys`,
    -`Start of productivity surveys`,
    -`End of occupancy surveys`,
    -`End of productivity surveys`,
    -`Number of productivity survey days`,
    -`Number of occupancy survey days`,
    -`Eggs`,	
    -`Eggs per occupied nest`,	
    # -`Young`,
    # -`Young per successful nest`,
    -`Fledglings`,	
    -`Fledglings per successful nest`,
    -`Productivity known`,

)
Summary_Current_char = sapply(Summary_Current, as.character) 
Summary_Current_char_DF = as.data.frame(Summary_Current_char)
colnames(Summary_Current_char_DF) = 'Result'

#Output
kable(Summary_Current_char_DF,col.names=c("Metric","Result"))

# Increment the table counter
TableCounter = TableCounter + 1

```

```{r echo=FALSE}

# Get the spatial locations of successful and failed nests for a map
Sql = paste0("SELECT   
Nest_ID
, RiverSegment
, Latitude AS Latitude
, Longitude AS Longitude
, NumYng
, Productivity
, CASE
	WHEN NOT Productivity IN ('Successful','Failed') THEN 'Not initated, not found, or productivity unknown'
	ELSE Productivity
END As Status
, StatusCode_P
, Nest_O
FROM Dataset_Productivity
WHERE (Year = ",SurveyYear,") 
AND (Latitude IS NOT NULL) -- Avoid missing spatial coords
AND (Latitude > 50) -- Avoid Lat/Lon records of 0/0
--AND (StatusCode_P IN ('S', 'IF')) -- Successful or failed, ignore not found and not counted records")

# Get data frame from query
Nests = dbGetQuery(Connection,Sql)
#Nests[is.na(Nests)] = " "

# Plot the nest success on a map
ggplot() +
  
  # Blue rectangle to make the background of the map blue
  geom_rect(aes(xmin = -155,xmax = -140,ymin = 55,ymax = 70),fill = "steelblue3") +
  
  # Alaska outline
  geom_sf(data=AK,mapping=aes(), color='darkgray',fill='snow2') +
  geom_sf(data=Parks,mapping=aes(), color='yellow4',fill='yellow4',alpha=0.3) +
  geom_sf(data=Rivers %>% filter(TYPE != 'I'),mapping=aes(),color='steelblue3',fill='steelblue3') +
  geom_sf(data = Roads, fill = "gray50", color = 'gray50',alpha=0.4) +
    
  # Failed nests
  # Consider carefully whether to publish nest locations, or not
  # geom_point(data=Nests, aes(x = Longitude, y = Latitude,colour=Productivity,shape=Productivity,size=Productivity)) +
  # scale_color_manual(values = c('Not initiated' = 'gray70','Successful' = 'green','Failed' = 'red','Productivity unknown' ='orange','Not found or not checked' = 'gray70')) +
  # scale_shape_manual(values = c('Not initiated' = 16,'Successful' = 16,'Failed' = 16,'Productivity unknown' = 16,'Not found or not checked' = 16)) +
  # scale_size_manual(values = c('Not initiated' = 1,'Successful' = 3,'Failed' = 2,'Productivity unknown' = 1,'Not found or not checked' = 1)) +
  
  geom_point(data=Nests, aes(x = Longitude, y = Latitude,colour=Status,shape=Status,size=Status)) +
  scale_color_manual(values = c('Not initated, not found, or productivity unknown' = 'black','Successful' = 'darkgreen','Failed' = 'red')) +
  scale_shape_manual(values = c('Not initated, not found, or productivity unknown' = 16,'Successful' = 16,'Failed' = 16)) +
  scale_size_manual(values = c('Not initated, not found, or productivity unknown' = 1,'Successful' = 3,'Failed' = 2)) +
  
  # Productive nests, increasing size with productivity
  # geom_point(data=subset(Nests,Productivity=="Successful"), aes(x = Longitude, y = Latitude, size=factor(NumYng)),color="black") +
  # # Override the size attributes assigned by geom_point. 'values' is size of symbol.
  # scale_size_manual(name="Young produced",values = c(2,5),breaks = c(1,2), labels = c("1","2")) + 
  
  # Set the map limits
  coord_sf(xlim = c(-146.6, -142), ylim = c(60.5, 62.8), expand = FALSE) +
  xlab('Longitude') +
  ylab('Latitude')+

  # WRST Label
  geom_text(aes(x = -143.5, y = 61.8, label = "Wrangell - St. Elias 
National Park and Preserve"), size = 3, color = "gray40") +
  
    # Upper Copper River label
  geom_text(aes(x = -144, y = 62.45, label = "Upper Copper\nRiver"), size = 3, color = "dodgerblue4") +
  
  # Middle Copper River label
  geom_text(aes(x = -145.2, y = 61.5, label = "Middle Copper River"), size = 3, color = "dodgerblue4") +
  
  # Lower Copper River label
  geom_text(aes(x = -145.2, y = 60.8, label = "Lower Copper River"), size = 3, color = "dodgerblue4") +
  
  # Chitina River label
  geom_text(aes(x = -143.8, y = 61.25, label = "Chitina River"), size = 3, color = "dodgerblue4") +
  
  # Bremner River label
  geom_text(aes(x = -143.8, y = 60.95, label = "Bremner River"), size = 3, color = "dodgerblue4") +
  
  # Minimal theme
  theme_minimal() +
  
  # Legend position
  theme(legend.position = "bottom")

```

Figure `r FigureCounter`. Bald eagle nest distribution and productivity in the Copper River Valley, `r SurveyYear`. Mean occupancy rate was `r PctOccupied`% and mean success rate was `r PctSuccessful`%. 


Bald eagle productivity metrics are presented by river segment in Table `r TableCounter`.

Table `r TableCounter`. Bald eagle productivity and success rates by river segment, `r SurveyYear`. MCR = Middle Copper River, LCR = Lower Copper River, UCR = Upper Copper River, BR = Bremner River, CHIT = Chitina River.

```{r echo=FALSE,label="Bald eagle productivity and success rates by river segment"}
# Increment the figure counter
FigureCounter = FigureCounter + 1

# Retrieve the annual summary information from the database
Sql = paste("SELECT Year
, [River segment]
,CASE
	WHEN [River segment] = 'Bremner' THEN 'BR'
	WHEN [River segment] = 'Chisana' THEN 'CHIS'
	WHEN [River segment] = 'Chitina' THEN 'CHIT'
	WHEN [River segment] = 'Copper' THEN 'CR'
	WHEN [River segment] = 'Gulkana' THEN 'GULK'
	WHEN [River segment] = 'Klutina' THEN 'KLUT'
	WHEN [River segment] = 'Lower Copper' THEN 'LCR'
	WHEN [River segment] = 'Middle Copper' THEN 'MCR'
	WHEN [River segment] = 'Ptarmigan' THEN 'PTAR'
	WHEN [River segment] = 'Tanana' THEN 'TAN'
	WHEN [River segment] = 'Tasnuna' THEN 'TAS'
	WHEN [River segment] = 'Tonsina' THEN 'TONS'
	WHEN [River segment] = 'Upper Copper' THEN 'UCR'
	WHEN [River segment] = 'Yakutat' THEN 'YAK'
END as [Segment]
, Records
, [Available nests]
, [Available but omitted]
, [Nests located]
, [Nests not located]
, [Nests not found]
, [Nests not counted]
, [Productivity unknown]
, [Productivity known]
, New
, Lost
, [Net gain]
, [Ending nest count]
, Occupied
, [Pct. occupied]
, Successful
, [Pct. successful]
, Eggs
, [Eggs per successful nest]
, Young
, [Young per successful nest]
, Fledglings
, [Fledglings per successful nest]
, [Start of occupancy surveys]
, [End of occupancy surveys]
, [Number of occupancy survey days]
, [Start of productivity surveys]
, [End of productivity surveys]
, [Number of productivity survey days]
, [Certified records]
FROM DashboardByRiverSegment
WHERE (Year <= ",SurveyYear,") 
AND ([River segment] <> 'Chitina River Tributaries') 
AND (NOT ([River segment] = 'Chitina')) OR (Year <= 2025) 
AND ([River segment] <> 'Chitina River Tributaries') AND (NOT (Year = 1989))
ORDER BY Year",sep="")

# Get the data from the database into a data frame
SummaryByRiverSegment = dbGetQuery(Connection,Sql) 

# Filter the above data frame on the current year and omit the first (Year) column
CurrentSummaryByRiverSegment = SummaryByRiverSegment %>%
  filter(Year == SurveyYear) %>%
  mutate(
    `Pct. occupied` = round(`Pct. occupied`,1),
    `Pct. successful` = round(`Pct. successful`,1),
    `Young per successful nest` = round(`Young per successful nest`,2)
  ) %>%
  select(`Segment`,
         -`Available nests`,
         -`Available but omitted`,
         `Nests located`,
         -`Nests not located`,
         -`Productivity unknown`,
         -`Productivity known`,
         New,
         Lost,
         `Net gain`,
         Occupied,
         Successful,
         `Pct. occupied`,
         `Pct. successful`,
         Young,
         `Young per successful nest`
)

# Output the data
kable(t(CurrentSummaryByRiverSegment))

# Increment the table counter
TableCounter = TableCounter + 1
```


```{r,echo=FALSE}
df_long <- CurrentSummaryByRiverSegment %>% 
  select(`Segment`,`Pct. occupied`,`Pct. successful`) %>% 
  #mutate(Failed = 100 - (PctOccupied + PctSuccessful)) %>%
  pivot_longer(
    cols = c(`Pct. occupied`,`Pct. successful`),
    names_to = "Category",
    values_to = "Percent"
  ) %>%
  mutate(
    Category = recode(Category,
                      `Pct. occupied` = "% Occupied",
                      `Pct. successful` = "% Successful")) %>%
  arrange(`Segment`,Category)

# Stacked + percent
ggplot(df_long, aes(fill = Category, y = Percent, x = `Segment`)) +
  geom_bar(stat = "identity", width = 0.6, position = position_dodge(width = .4)) +
  scale_fill_grey(start = 0.2, end = 0.6) +
  theme_grey() +
  theme(legend.position = "bottom")


```

Figure `r FigureCounter`. Percent nest occupancy and success by river segment, `r SurveyYear`. MCR = Middle Copper River, LCR = Lower Copper River, UCR = Upper Copper River, BR = Bremner River, CHIT = Chitina River.


```{r echo=FALSE,include=FALSE}
if(1==0){
  # This chunk copies the report to the J or O drive
  CurrentDirectory = "C:/Development/WRST BAEA Reports R Markdown/"
  From = paste(CurrentDirectory,"WRST BAEA Annual Summary Report Short.pdf",sep='')
  To = paste(CurrentDirectory,SurveyYear," WRST BAEA Survey Summary.pdf",sep='') # ARCN
  cat(From,' to ',To)
  
  # Copy the file to the Network reports directory
  file.copy(From, To, overwrite = TRUE)
}

```

<!-- ## R Markdown Report Template -->

<!-- This template was generated automatically from an R Markdown script available at <https://github.com/NPS-ARCN-CAKN/WRST-BAEA-Annual-Report-R-Markdown>. The information in the report was generated from bald eagle survey data in the NPS WRST_Caribou SQL Server database (inpyugamsvm01\nuna:WRST_Caribou). -->
