---
title: ""
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

```{=html}
<style>

body {
    font-family: "Times New Roman";
    font-size: 11.5pt;
    color: black;
}

.title {
  font-family: "Times New Roman";
  font-size: 16pt;
  font-weight: bold;
  color: black;
}

h1 {
    font-family: "Arial";
    font-size: 16pt;
    font-weight: bold;
    color: black;
}


h2 {
  font-family: "Arial";
  font-size: 11.5pt;
  font-weight: bold;
  color: black;
}

h3 {
  font-family: "Arial";
  font-size: 10.5pt;
  font-weight: bold;
  color: black;
}

h4 {
    font-family: "Times New Roman";
    font-size: 12pt;
    color: black;
}

p {
    font-family: "Times New Roman";
    font-size: 11.5pt;
    color: black;
}

.caption {
    font-family: "Arial";
    font-size: 10pt;
    color: black;
}

th {
    font-family: "Arial";
    font-size: 9pt;
    font-weight: bold;
    color: black;
}

tr,td,table {
    font-family: "Arial";
    font-size: 9pt;
    color: black;
}

.header{
  font-family: "Arial";
  font-size: 10pt;
  font-weight: bold;
  color: black;
}

</style>
```

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) # Prevent overwhelming messages from reaching user

options(download.file.method = "wininet")
#install.packages("sqldf")
# Load libraries
library(RODBC)
library(knitr)
library(ggplot2)
library(sp)
library(sf)
library(sqldf)
library(tidyverse)


# Make kable show blanks for NAs
options(knitr.kable.NA = '')

# These parameters restrict the dimensions of plots to 6 metric inches so they don't blow out Word
FigWidth = 6
```

```{r echo=FALSE,label="SET UP REPORT PARAMETERS HERE"}

# Set up report parameters here
SurveyYear = 2024  # Set this to the desired report year. Don't set to 2016 because that survey didn't come off and the script will bomb

# ODBC connection to the BAEA monitoring SQL Server Database. 
# Contact data manager for server and database details.
# ConnectionString to the bald eagles Sql Server database
BaldEaglesConnectionString = "Driver=SQL Server Native Client 11.0;Server=inpyugamsvm01\\nuna;Database=CAKN_BaldEagles;trusted_connection=YES"

# Database connection to the eagles database
BaldEaglesConnection = odbcDriverConnect(connection = BaldEaglesConnectionString)

# Set the path to shapefiles used in the map, AK, Parks, Rivers, etc
ShpPath = "C:/Work/GIS Common Layers/"

```

---
title: Annual Report on Vital Signs monitoring of distribution and abundance of Bald Eagles in Wrangell-St. Elias National Park and Preserve, `r SurveyYear` 
output:
  html_document:
---

```{r echo=FALSE,label="Functions"}

# Current date
CurrentDate = format(Sys.Date(), format='%B %d, %Y') # Get the current date

# Function to return data frames from SQL database queries
GetDataFrameFromDatabase <- function(Sql) {
   return(sqlQuery(BaldEaglesConnection, Sql))
}

# Function to convert a date string to a long date
GetLongDate <- function(Date) {
  return(format(as.Date(Date), format="%B %d, %Y"))
}

# These counters are used to number the tables and figures. Each gets incremented after use.
TableCounter = 1
FigureCounter = 1

```

```{r echo=FALSE, label="Survey dates"}

# Summary_SurveyDatesByYearDataFrame: Survey Dates
Sql = paste("SELECT 
--[Year],
[MinOccupancySurveyDate] as [Occupancy survey start]
,[MaxOccupancySurveyDate] as [Occupancy survey end]
-- ,[NumberOfOccupancySurveyDays]
,[MinProductivitySurveyDate] as [Productivity survey start]
,[MaxProductivitySurveyDate] as [Productivity survey end]
-- ,[NumberOfProductivitySurveyDays]
FROM [CAKN_BaldEagles].[dbo].[Summary_SurveyDatesByYear] 
WHERE [Year] = ",SurveyYear,sep="")
Summary_SurveyDatesByYearDataFrame = GetDataFrameFromDatabase(Sql) # Get the data from the database into a data frame

# Build some survey date values to use later in the markdown text
OStart = Summary_SurveyDatesByYearDataFrame$`Occupancy survey start`
OEnd = Summary_SurveyDatesByYearDataFrame$`Occupancy survey end`
PStart = Summary_SurveyDatesByYearDataFrame$`Productivity survey start`
PEnd = Summary_SurveyDatesByYearDataFrame$`Productivity survey end`

# Summary_SuccessByYear: Success and productivity data by year
Sql = paste("SELECT [Year]
,[Nests located]
,[Occupied]
,[Successful]
,[Productivity unknown]
--,[Eggs]
,[Young fledged]
,[% nest occupied]
,[% nest success]
,[Young/occupied nest]
,[Young/successful nest]
--,Source
--,Left(Source,4) as Source
FROM [Summary_SuccessByYear]
WHERE Year <= ",SurveyYear," 
AND Year NOT IN (2016) -- 2016 Interrupted by bad weather
ORDER BY [Year]",sep="")

# Get the data from the database into a data frame
Summary_SuccessByYearDataFrame = GetDataFrameFromDatabase(Sql) 

# Length of record
MinYear = min(Summary_SuccessByYearDataFrame$Year, na.rm = TRUE) 
MaxYear = max(Summary_SuccessByYearDataFrame$Year, na.rm = TRUE)
LengthOfRecord = MaxYear - MinYear
FirstSurveyYear = min(Summary_SuccessByYearDataFrame$Year)
NumberOfSurveys = sqldf('SELECT Count(DISTINCT Year) as NumberOfSurveys FROM Summary_SuccessByYearDataFrame')

# We'll need just this year's data in a separated data frame from all the historical data.
# Filter on the summary data frame on the current year and then remove the Year column
Summary_SuccessByYearDataFrame_Current = Summary_SuccessByYearDataFrame[Summary_SuccessByYearDataFrame$Year == SurveyYear,-1]

# Build some summary statistics to report in the markdown
NestsLocated = Summary_SuccessByYearDataFrame_Current$`Nests located`
Occupied = Summary_SuccessByYearDataFrame_Current$Occupied
Successful = Summary_SuccessByYearDataFrame_Current$Successful
YoungFledged = Summary_SuccessByYearDataFrame_Current$`Young fledged`
NF = Summary_SuccessByYearDataFrame_Current$`Occupied nests not found during productivity survey)`
NC = Summary_SuccessByYearDataFrame_Current$`Occupied nests not counted during productivity survey)`
YngPerOccupiedNest = signif(Summary_SuccessByYearDataFrame_Current$`Young/occupied nest`,3)
YngPerSuccessfulNest = signif(Summary_SuccessByYearDataFrame_Current$`Young/successful nest`,3)
PctOccupied = signif(Summary_SuccessByYearDataFrame_Current$`% nest occupied`,2)
PctSuccessful = signif(Summary_SuccessByYearDataFrame_Current$`% nest success`,2)
ProductivityUnknown = Summary_SuccessByYearDataFrame_Current$`Productivity unknown`

ExecutiveSummaryText = paste("The Copper River in Wrangell-St. Elias National Park and Preserve (WRST), Alaska, contains a high density of nesting Bald Eagles <i>(Haliaeetus leucocephalus)</i> and the National Park Service has identified Bald Eagles as a focal fauna species to monitor. Bald Eagles in WRST are a high profile species that are dependent upon many resources along the Copper River and nest at the northern edge of the their range. Bald Eagles are top-trophic level predators and they often respond quickly to changes in their environment by changing their breeding activities. In ",SurveyYear," we conducted an aerial bald eagle productivity survey in and around the park. Nest occupancy surveys were flown between ",GetLongDate(OStart)," and ",GetLongDate(OEnd),", followed by productivity surveys between ",GetLongDate(PStart)," and ",GetLongDate(PEnd),". We observed ",NestsLocated," nests, of which ",Occupied," were occupied (",PctOccupied,"%). Productivity surveys revealed that ",Successful," occupied nests produced young (",PctSuccessful,"% of occupied nests. Productivity could not be determined for ",ProductivityUnknown," nests). Successful nests resulted in ",YoungFledged," young, an average of ",YngPerSuccessfulNest," young per successful nest.",sep="")


```

Kyle Cutting^1^ and Scott Miller^2^

^1^National Park Service, Wrangell St. Elias National Park and Preserve, 8 Richardson Hwy Mile 106, Copper Center, AK 99573

^2^National Park Service, Arctic Inventory and Monitoring Network,240 W. 5th Ave Anchorage, Alaska 99501

`r CurrentDate`

U.S. Department of the Interior\
National Park Service\
Natural Resource Stewardship and Science\
Fort Collins, Colorado

# Abstract

`r ExecutiveSummaryText`

**Keywords:** Wrangell-St. Elias National Park and Preserve, Bald Eagle, *Haliaeetus leucocephalus*, Population dynamics, monitoring

# Objectives

The objectives of the monitoring efforts are to:

-   Document the annual changes in occupancy of bald eagles,

-   Document the annual changes in productivity of bald eagle nests, and

-   Document any changes in distribution of bald eagle nests.

# Methods

The `r SurveyYear` bald eagle survey followed the same methods (Putera and Miller, 2018) that have been used since `r MinYear`. Five survey areas were defined along the Copper River and its major tributaries (Figure 1):

1.) The Upper Copper River (UCR, 80km) extends from Copper and Tanada lakes to the confluence of the Chistochina River. White spruce (*Picea glauca*) is the predominant tree species, with balsam poplar (*Populus balsamifera*) common on the lower stretches.

2.) The Middle Copper River (MCR, 179km) extends from the Chistochina River south to the confluence of the Chitina River.The MCR contains a mixture of both white spruce and balsam poplar, with poplar being slightly more common. The Copper River begins to braid along the MCR.

3.) The Lower Copper River (LCR) section (188km) extends from the Chitina River south to Miles Lake. The LCR is dominated by cottonwood along the river. The vegetation also begins to exhibit coastal influences, adding several coastal species including black cottonwood (*Populus trichocarpa*), Sitka spruce (*Picea sitchensis*), and mountain hemlock (*Tsuga mertensiana*). The Copper River is wide and more braided than the upper sections, with steep canyons on either side. Nesting begins consistently later along the LCR and is probably influenced by the numerous glaciers in the area which delay snow melt and maintain the area in a winter-like state several weeks after green up has begun to occur to the north.

4.) The Bremner River (45km) is surveyed from its confluence with the Copper River to the forks by Threemile Canyon, and has similar characteristics as the LCR.

5.) The Chitina River (116km) is surveyed from the confluence of the Copper River to where the river becomes braided above the Tana River. White spruce dominate the trees, with cottonwood becoming scarce in the upper reaches.

```{r echo=FALSE,label="Survey area map",fig.alt=paste("Figure ",FigureCounter,". Map of the bald eagles monitoring survey area in the Copper River drainage, Wrangell-St. Elias National Park and Preserve. Major tributaries in the survey area include the Chitina and Bremner rivers.",sep=""),fig.cap=paste("Figure ",FigureCounter,". Map of the bald eagles monitoring survey area in the Copper River drainage, Wrangell-St. Elias National Park and Preserve. Major tributaries in the survey area include the Chitina and Bremner rivers.",sep=""),fig.width=FigWidth}

# Load shapefiles to be used in the maps later.
Parks = read_sf(paste(ShpPath,"GCS/AKParks_reprojected.shp",sep=""))
Rivers = read_sf(paste(ShpPath,"Alaska_Rivers_1M.shp",sep=""))
AK  = read_sf(paste(ShpPath,"GCS/AK_GCS_poly.shp",sep=""))

# Build a map of the survey area showing park boundaries, copper river and tributaries, etc.
ggplot() +
  
  # Blue rectangle to make the background of the map blue
  geom_rect(aes(xmin = -155,xmax = -140,ymin = 55,ymax = 70),fill = "steelblue3") +
  
  # Alaska outline
  geom_sf(data=AK,mapping=aes(), color='darkgray',fill='snow2') +
  
  # Park boundary
  geom_sf(data=Parks,mapping=aes(), color='yellow4',fill='yellow4',alpha=0.3) +
  
  # Rivers
  geom_sf(data=Rivers,mapping=aes(),color='steelblue3',fill='steelblue3') +
  
  # Set the map limits, zoom into the BAEA monitoring area
  coord_sf(xlim = c(-146.6, -142), ylim = c(60.5, 62.8), expand = FALSE) +
  xlab('Longitude') +
  ylab('Latitude')+
  
  # Minimal theme
  theme_minimal() +
  
  # WRST text Label
  geom_text(aes(x = -144.25, y = 61.8, label = "Wrangell - St. Elias National\nPark and Preserve"), size = 4, color = "gray40", fontface = "bold") +
  
  # Upper Copper River label
  geom_text(aes(x = -144, y = 62.45, label = "Upper Copper River"), size = 4, color = "black") +
  
  # Middle Copper River label
  geom_text(aes(x = -145.2, y = 61.5, label = "Middle Copper River"), size = 4, color = "black") +
  
  # Lower Copper River label
  geom_text(aes(x = -145.2, y = 60.8, label = "Lower Copper River"), size = 4, color = "black") +
  
  # Chitina River label
  geom_text(aes(x = -143.8, y = 61.25, label = "Chitina River"), size = 4, color = "black") +
  
  # Bremner River label
  geom_text(aes(x = -143.8, y = 60.95, label = "Bremner River"), size = 4, color = "black")

```

Two aerial surveys were conducted in `r SurveyYear` - one to document territorial occupancy, and one to document productivity. Surveys were flown using a small airplane flying at low level (\<100m AGL) and low speed (\<100 kph), with one experienced observer. Flight paths followed the river's edge, with the opposite side covered on the return trip. In most cases, the relatively narrow riparian corridor allowed adequate coverage with a single pass on each side of the river. Where the corridor widened, multiple passes were made to adequately evaluate the available habitat.

For all new nests located, the following information was collected:

a)  Location (latitude/longitude via handheld Global Positioning System receiver),
b)  Count of adult birds,
c)  Behavior of adults (incubating, perched),
d)  Number of eggs observed,
e)  Number of nestlings,
f)  Species of nest tree (Populus spp., Picea spp.),
g)  Whether the nest tree was alive or dead, and
h)  Nest status (good, poor, remnant, gone).

All information except latitude/longitude was recorded on field forms. All nests still intact from the previous year's surveys were checked for occupancy during the occupancy survey. Information recorded for located nests include:

a)  Count of adult birds,
b)  Behavior of adults, and
c)  Number of eggs or nestlings observed.

During the second survey, all nest locations collected during prior surveys were checked for productivity. Territorial occupancy was defined based on a bird in the nest in incubating posture or two adult birds at a nest or in the nest tree (Postupalsky, 1974). For the productivity survey, all nests observed with an incubating bird or two adult birds at a nest during the occupancy survey were checked to determine nest outcome. Multiple passes were commonly needed to count chicks, particularly in the Lower Copper area where cottonwoods are more common, and leaves make counting chicks difficult.

# Results

```{r echo=FALSE,label="Get metrics into variables to use in markdown text"}

# Build a paragraph of information about the survey results into a variable. Embed this variable in the markdown text later (the stats could be embedded in the markdown individually, but for some reason it doesn't always work consistently, so I built the paragraph here.)
ResultsText = paste('Overall, ',NestsLocated, ' nests were located during the occupancy survey flights (Table 1). Of these nests, ',Occupied,' were occupied (', PctOccupied,'%) and of occupied nests, ',Successful,' successfully produced young (',PctSuccessful,'%, productivity could not be determined for ',ProductivityUnknown,' nests). ', str_to_sentence(xfun::numbers_to_words(YoungFledged)),' young were produced overall. Occupied nest production averaged ', YngPerOccupiedNest,' young per nest and successful nests produced a mean of ',YngPerSuccessfulNest,' young.',sep='')

```

Bald eagle occupancy survey flights were conducted in WRST from `r GetLongDate(OStart)` to `r GetLongDate(OEnd)`. Follow-up productivity flights happened from `r GetLongDate(PStart)` to `r GetLongDate(PEnd)`. `r ResultsText` Table `r TableCounter` shows nest success and productivity metrics by river segment. Spatial distribution of surveyed nests and their productivity is shown in Figure `r FigureCounter`.

```{r echo=FALSE, label="Current survey metrics table"}
# Summarize the eagle counts for the current survey into 
# a table, transpose it so it fits the report better, modify the header, and output

# Convert it all to strings so R doesn't add decimal places to integers
Summary_SuccessByYearDataFrame_Current = sqldf(paste("SELECT 
[Nests located]
,Occupied
,Successful
,[Productivity unknown]
,[Young fledged]
,[% nest occupied]
,[% nest success]
,[Young/occupied nest] 
,[Young/successful nest] 
FROM Summary_SuccessByYearDataFrame WHERE Year = ",SurveyYear,sep=''))
Summary_SuccessByYearDataFrame_Current_char = sapply(Summary_SuccessByYearDataFrame_Current, as.character) 

# Convert it into a data frame
Summary_SuccessByYearDataFrame_Current_char_DF = as.data.frame(Summary_SuccessByYearDataFrame_Current_char)
colnames(Summary_SuccessByYearDataFrame_Current_char_DF) = 'Result'

#Output
kable(Summary_SuccessByYearDataFrame_Current_char_DF,col.names=c("Metric","Result"), caption=paste("Table ",TableCounter,". Results summary, ",SurveyYear," bald eagles survey.",sep=""))

# Increment the table counter
TableCounter = TableCounter + 1

# Set up a caption for the next figure
FigureCaption = paste("Figure ",FigureCounter,". Bald eagle nest dispersion and breeding pair productivity metrics in the Copper River Valley in ",SurveyYear,". Black circles represent successful nests with the size indicating number of young produced. Failed nests are shown with an X. Mean occupancy rate was ",PctOccupied,"% and mean success rate was ",PctSuccessful,"%.",sep='')
```

```{r echo=FALSE, label="Success map",fig.alt=FigureCaption,fig.cap=FigureCaption,fig.width=FigWidth}

# Get the spatial locations of successful and failed nests for a map
Sql = paste("SELECT   
--Nest_ID
--, RiverSegment
[Lat-NAD83] AS Latitude
, [Lon-NAD83] AS Longitude
, NumYng
,CASE 
  WHEN StatusCode_P = 'IF' THEN 'Failed' 
  WHEN StatusCode_P = 'S' THEN 'Succeeded' 
  ELSE NULL 
END AS Status
, StatusCode_P
--, Nest_O
FROM Dataset_Productivity
WHERE (Year = ",SurveyYear,") 
AND ([Lat-NAD83] IS NOT NULL) -- Avoid missing spatial coords
AND ([Lat-NAD83] > 50) -- Avoid Lat/Lon records of 0/0
AND (StatusCode_P IN ('S', 'IF')) -- Successful or faild, ignore not found and not counted records
",sep='')

# Get data frame from query
Dataset_ProductivityDataFrame = GetDataFrameFromDatabase(Sql)
Dataset_ProductivityDataFrame[is.na(Dataset_ProductivityDataFrame)] = " "

# Map Title
title = paste("Dispersion of successful nests, ", SurveyYear)

# Query out just the successful nests
#SuccessFulNestsDataFrame = subset(Dataset_ProductivityDataFrame,StatusCode_P == 'S')

# Plot the nest success on a map
ggplot() +
  
  # Blue rectangle to make the background of the map blue
  geom_rect(aes(xmin = -155,xmax = -140,ymin = 55,ymax = 70),fill = "steelblue3") +
  
  # Alaska outline
  geom_sf(data=AK,mapping=aes(), color='darkgray',fill='snow2') +
  
  # Park boundary
  geom_sf(data=Parks,mapping=aes(), color='yellow4',fill='yellow4',alpha=0.3) +

    # Rivers
  geom_sf(data=Rivers,mapping=aes(),color='steelblue3',fill='steelblue3') +  
  
  # Failed nests
  # Consider carefully whether to publish nest locations, or not
  geom_point(data=Dataset_ProductivityDataFrame, aes(x = Longitude, y = Latitude,colour=Status,na.rm = TRUE,shape=Status),size=2) +
  scale_color_manual(values = c("Succeeded" = "black", "Failed" = "black")) +
  scale_shape_manual(values = c(4,19)) + # 4 = X, 19 = dot
  
  # Productive nests, increasing size with productivity
  geom_point(data=subset(Dataset_ProductivityDataFrame,Status=="Succeeded"), aes(x = Longitude, y = Latitude, size=factor(NumYng), na.rm = TRUE),color="black") +
  # Override the size attributes assigned by geom_point. 'values' is size of symbol.
  scale_size_manual(name="Young produced",values = c(2,5),breaks = c(1,2), labels = c("1","2")) + 
  
  # Set the map limits
  coord_sf(xlim = c(-146.6, -142), ylim = c(60.5, 62.8), expand = FALSE) +
  xlab('Longitude') +
  ylab('Latitude')+
  
  # Minimal theme
  theme_minimal() +
  
  # WRST Label
  geom_text(aes(x = -143.5, y = 61.8, label = "Wrangell - St. Elias 
National Park and Preserve"), size = 3, color = "gray40") +
  
    # Upper Copper River label
  geom_text(aes(x = -144, y = 62.45, label = "Upper Copper\nRiver"), size = 3, color = "dodgerblue4") +
  
  # Middle Copper River label
  geom_text(aes(x = -145.2, y = 61.5, label = "Middle Copper River"), size = 3, color = "dodgerblue4") +
  
  # Lower Copper River label
  geom_text(aes(x = -145.2, y = 60.8, label = "Lower Copper River"), size = 3, color = "dodgerblue4") +
  
  # Chitina River label
  geom_text(aes(x = -143.8, y = 61.25, label = "Chitina River"), size = 3, color = "dodgerblue4") +
  
  # Bremner River label
  geom_text(aes(x = -143.8, y = 60.95, label = "Bremner River"), size = 3, color = "dodgerblue4")

# Increment the figure counter
FigureCounter = FigureCounter + 1
```

Bald eagle productivity metrics are presented by river segment in Table `r TableCounter`.

```{r echo=FALSE,label="Bald eagle productivity and success rates by river segment"}

# Retrieve the annual summary information from the database
Sql = paste("SELECT  [Year]
,[RiverSegment] as [River segment]
,[Nests located]
,[Occupied]
,[Successful]
,[Productivity unknown]
--,[Eggs] -- Eggs were inconsistently collected over the years
,[Young fledged]
,[% nest occupied]
,[% nest success]
,[Young/occupied nest]
,[Young/successful nest]
,[Occupied nests not found during productivity survey)]
,[Occupied nests not counted during productivity survey)],Source
,Notes
FROM [CAKN_BaldEagles].[dbo].[Summary_SuccessByYearAndRiverSegment]
WHERE Year <= ",SurveyYear," 
And RiverSegment <> 'Chitina River Tributaries' -- only done once
And NOT(RiverSegment = 'Chitina' And Year = 1989) -- Only two nests observed, one successful, inflates chitina productivity so omit
ORDER BY Year",sep="")

# Get the data from the database into a data frame
SuccessByYearRiverSegmentDF = GetDataFrameFromDatabase(Sql) 

# Filter the above data frame on the current year and omit the first (Year) column
CurrentSuccessByYearRiverSegmentDF = SuccessByYearRiverSegmentDF[SuccessByYearRiverSegmentDF$Year == SurveyYear,2:11]

# Change all the values to strings so R does not pad zeroes to the integer values
CurrentSuccessByYearRiverSegmentDF_str = sapply(CurrentSuccessByYearRiverSegmentDF, as.character) 

# Transpose the data
CurrentSuccessByYearRiverSegmentDF_str_t = t(CurrentSuccessByYearRiverSegmentDF_str)
colnames(CurrentSuccessByYearRiverSegmentDF_str_t)[1:5] = "" # Change the wierd first column with no name that R stuck in to Result

# Output the data
kable(CurrentSuccessByYearRiverSegmentDF_str_t, caption = paste("Table ",TableCounter,". Bald eagle productivity and success rates by river segment, ",SurveyYear,".",sep=""))

# Increment the table counter
TableCounter = TableCounter + 1
```

## Long term trends

```{r echo=FALSE,label="Build long term success and productivity metrics over the entire surveying history to use in the markdown"}

# Build long term success and productivity metrics over the entire surveying history to use in the markdown

# Occupancy rate %
MeanOccupancy = signif(round(mean(Summary_SuccessByYearDataFrame$`% nest occupied`, na.rm = TRUE),1),2)
MinOccupancy= signif(min(Summary_SuccessByYearDataFrame$`% nest occupied`, na.rm = TRUE),2)
MaxOccupancy= signif(max(Summary_SuccessByYearDataFrame$`% nest occupied`, na.rm = TRUE),2)

# Success rate %
MeanSuccess = signif(mean(Summary_SuccessByYearDataFrame$`% nest success`, na.rm = TRUE),2)
MinSuccess= signif(min(Summary_SuccessByYearDataFrame$`% nest success`, na.rm = TRUE),2)
MaxSuccess= signif(max(Summary_SuccessByYearDataFrame$`% nest success`, na.rm = TRUE),2)

# Productivity per nest
MeanYoungPerOccupiedNest = signif(mean(Summary_SuccessByYearDataFrame$`Young/occupied nest`, na.rm = TRUE),3)
MinYoungPerOccupiedNest = signif(min(Summary_SuccessByYearDataFrame$`Young/occupied nest`, na.rm = TRUE),3)
MaxYoungPerOccupiedNest = signif(max(Summary_SuccessByYearDataFrame$`Young/occupied nest`, na.rm = TRUE),3)
MeanYoungPerSuccessfulNest = signif(mean(Summary_SuccessByYearDataFrame$`Young/successful nest`, na.rm = TRUE),3)
MinYoungPerSuccessfulNest = signif(min(Summary_SuccessByYearDataFrame$`Young/successful nest`, na.rm = TRUE),3)
MaxYoungPerSuccessfulNest = signif(max(Summary_SuccessByYearDataFrame$`Young/successful nest`, na.rm = TRUE),3)

# Build a long term results summary paragraph to use in the markdown below
ResultsText = paste("Over the ",LengthOfRecord," year history of the bald eagles monitoring effort in the Copper River study area, there have been ",NumberOfSurveys," surveys (Table ",TableCounter,", Figures 2 and 3). During this time, mean nest occupancy has ranged from ", MinOccupancy,"% to ", MaxOccupancy,"%, averaging ",MeanOccupancy,"%. Success rates, defined as occupied nests producing young, has ranged from ",MinSuccess,"% to ",MaxSuccess,"%, averaging ",MeanSuccess,"%. Productivity averaged ",MeanYoungPerSuccessfulNest," young per successful nest, ranging between ",MinYoungPerSuccessfulNest," to ", MaxYoungPerSuccessfulNest,".",sep='')

```

`r ResultsText`

```{r echo=FALSE,label="Long-term bald eagle productivity metrics, Copper River Valley"}

# Show the long-term success and productivity metrics in a table
kable(Summary_SuccessByYearDataFrame, caption = paste("Table ",TableCounter,". Long-term bald eagle productivity metrics, Copper River Valley.",sep=""))

# Increment the table counter
TableCounter = TableCounter + 1

```

Notes: The 2016 survey was not finished due to bad weather and has been omitted from the results table. The 2009 productivity survey was hampered by smoke leading to a higher than average number of nests with unknown productivity.

## Success

Long-term bald eagle nest success metrics at WRST are shown in Figure `r FigureCounter`. Results for the years 1989 through 1996 are from Kozie (1996). Results from 2004 to present are from NPS Inventory and Monitoring survey data.

```{r echo = FALSE,label="Success rates plot prep"}
# This chunk generates a figure caption and alt text for the plot in the next R chunk. It has to be here in its own R chunk because it has to be ready to assign to the fig.alt and fig.cap attributes in the header of the next R chunk.

# Generate a linear model of success rate by year
# lm = lm(formula = `% nest success` ~ Year, data = Summary_SuccessByYearDataFrame)
# 
# # Summarize the linear model
# Summary = summary(lm)
# 
# # Extract linear model parameters from the Summary object. 
# R2 = signif(Summary$adj.r.squared, 3)
# AdjR2 = signif(Summary$adj.r.squared, 3)
# P = signif(Summary$coef[2,4],5)
# Intercept = signif(lm$coef[[1]],5 )
# Slope = signif(lm$coef[[2]],5)
# 
# # Make a linear model text summary to put in the figure caption for the plot of success by year
# lmStats_Success = paste("Adj R2 = ",AdjR2,"Intercept = ",Intercept, " Slope =",Slope, " P =",P)
# 
# # Generate a linear model of occupancy rate by year
# lm = lm(formula = `% nest occupied` ~ Year, data = Summary_SuccessByYearDataFrame)
# 
# # Summarize the linear model
# Summary = summary(lm)
# 
# # Extract linear model parameters from the Summary object. 
# R2 = signif(Summary$adj.r.squared, 3)
# AdjR2 = signif(Summary$adj.r.squared, 3)
# P = signif(Summary$coef[2,4],5)
# Intercept = signif(lm$coef[[1]],5 )
# Slope = signif(lm$coef[[2]],5)
# 
# # Make a linear model text summary to put in the figure caption for the plot of success by year
# lmStats_Occupancy = paste("Adj R2 = ",AdjR2,"Intercept = ",Intercept, " Slope =",Slope, " P =",P)

# Set up the figure caption for the next figure
FigureCaption = paste("Figure ",FigureCounter,". Nest success by year over the Copper River study area. Solid line shows mean nest occupancy rate (mean = ",MeanOccupancy,"%) with solid black smoothing line applied over it. Dotted line shows mean success rate for all occupied nests (mean = ",MeanSuccess,"%) with solid gray smoothing line applied over it.",sep='')

```

```{r echo=FALSE,label="Percent nest success by year",fig.alt=FigureCaption,fig.cap=FigureCaption,fig.width=FigWidth}

# Plot historical occupancy and success rates for the whole watershed
ggplot() +

  # Smooth lines
  geom_smooth(data=Summary_SuccessByYearDataFrame,aes(x=Year,y=`% nest success`),se=FALSE,color='black',size=0.5,method='loess') +
  geom_smooth(data=Summary_SuccessByYearDataFrame,aes(x=Year,y=`% nest occupied`),se=FALSE,color='gray',size=0.5,method='loess') +
  
  # % nest success curve
  # Kozie's data
  geom_line(data=subset(Summary_SuccessByYearDataFrame,Year <= 1997),mapping = aes(x=Year,y=`% nest success`,linetype="Percent success")) +
  # I&M data
  geom_line(data=subset(Summary_SuccessByYearDataFrame,Year > 1997),mapping = aes(x=Year,y=`% nest success`,linetype="Percent success")) +

  # % nest occupied curve (there is no Kozie, pre-2004 data for this)
  geom_line(data=subset(Summary_SuccessByYearDataFrame,Year > 1997),mapping = aes(x=Year,y=`% nest occupied`,linetype="Percent occupancy")) +

  # Fix the legend, define the line types
  scale_linetype_manual(name="",values = c("Percent success" = "solid","Percent occupancy" = "dotted"),) +
  
  # Label axes
  xlab("Year") +
  ylab("Occupancy and success rates") +
  
  # Y limits
  ylim(0,100) +
  
   # Minimal theme
  theme_minimal() +
  theme(legend.position = "top") 
  
# Increment the figure counter
FigureCounter = FigureCounter + 1

```

## Productivity

Long-term bald eagle nest productivity metrics at WRST are shown in Figure `r FigureCounter`. Results for the years 1989 through 1996 are from Kozie (1996). Results from 2004 to present are from NPS Inventory and Monitoring survey data.

```{r echo = FALSE,label="Productivity plot prep"}
# This chunk generates a figure caption and alt text for the plot in the next R chunk. It has to be here in its own R chunk because it has to be ready to assign to the fig.alt and fig.cap attributes in the header of the next R chunk.

# Generate a linear model of success rate by year
# lm = lm(formula = `Young/successful nest` ~ Year, data = Summary_SuccessByYearDataFrame)
# 
# # Summarize the linear model
# Summary = summary(lm)
# 
# # Extract linear model parameters from the Summary object. 
# R2 = signif(Summary$adj.r.squared, 3)
# AdjR2 = signif(Summary$adj.r.squared, 3)
# P = signif(Summary$coef[2,4],5)
# Intercept = signif(lm$coef[[1]],5 )
# Slope = signif(lm$coef[[2]],5)
# 
# # Make a linear model text summary to put in the figure caption for the plot of success by year
# lmStats_ProductivitySuccessful = paste("Adj R2 = ",AdjR2,"Intercept = ",Intercept, " Slope =",Slope, " P =",P)
# 
# # Generate a linear model of occupancy rate by year
# lm = lm(formula = `Young/occupied nest` ~ Year, data = Summary_SuccessByYearDataFrame)
# 
# # Summarize the linear model
# Summary = summary(lm)
# 
# # Extract linear model parameters from the Summary object. 
# R2 = signif(Summary$adj.r.squared, 3)
# AdjR2 = signif(Summary$adj.r.squared, 3)
# P = signif(Summary$coef[2,4],5)
# Intercept = signif(lm$coef[[1]],5 )
# Slope = signif(lm$coef[[2]],5)
# 
# # Make a linear model text summary to put in the figure caption for the plot of success by year
# lmStats_ProductivityOccupied = paste("Adj R2 = ",AdjR2,"Intercept = ",Intercept, " Slope =",Slope, " P =",P)

# Set up the figure caption for the next figure
FigureCaption = paste("Figure ",FigureCounter,". Bald eagle productivity by year over the Copper River study area. Solid black line shows mean productivity for successful nests (occupied and productive) with a solid black smoothing line applied over it. Mean long term productivity for successful nests is ",MeanYoungPerSuccessfulNest," young per nest). Dotted line shows mean productivity for all occupied nests (mean = ",MeanYoungPerOccupiedNest,") with a solid gray smoothing line applied over it.",sep='')
```

```{r echo=FALSE,label="Productivity by year plot",fig.alt=FigureCaption,fig.cap=FigureCaption,fig.width=FigWidth}

# Plot productivity
ggplot(data=Summary_SuccessByYearDataFrame) +
  
  # Smooth lines
  geom_smooth(data=Summary_SuccessByYearDataFrame,aes(x=Year,y=`Young/successful nest`),se=FALSE,color='black',size=0.5,method='loess') +
  geom_smooth(data=Summary_SuccessByYearDataFrame,aes(x=Year,y=`Young/occupied nest`),se=FALSE,color='gray',size=0.5,method='loess') +
  
  # Kozie's 1989-1996 data
  geom_line(data=subset(Summary_SuccessByYearDataFrame,Year <= 1997),mapping = aes(x=Year,y=`Young/successful nest`,linetype="Young/successful nest")) +
  
  # I&M data
  geom_line(data=subset(Summary_SuccessByYearDataFrame,Year > 1997),mapping = aes(x=Year,y=`Young/successful nest`,linetype="Young/successful nest")) +
  
  # Kozie's 1989-1996 data
  geom_line(data=subset(Summary_SuccessByYearDataFrame,Year <= 1997),mapping = aes(x=Year,y=`Young/occupied nest`,linetype="Young/occupied nest")) +
  
  #I&M data
  geom_line(data=subset(Summary_SuccessByYearDataFrame,Year > 1997),mapping = aes(x=Year,y=`Young/occupied nest`,linetype="Young/occupied nest")) +

  # Fix the legend, define the line types
  scale_linetype_manual(name="",values = c("Young/occupied nest" = "dotted","Young/successful nest" = "solid"),) +
  
  # Label axes
  xlab("Year") +
  ylab("Young per nest") +
  
  # Minimal theme
  theme_minimal() +
  theme(legend.position = "top")

# Increment the figure counter
FigureCounter = FigureCounter + 1

```

```{r , echo=FALSE,fig.height=10}
# Success by river segment plot
# ggplot(df) +
#   geom_line(aes(x=Year,y=`% nest success`,linetype=Source),na.rm=TRUE) +
#   geom_smooth(aes(x=Year,y=`% nest success`,color=Source),se=FALSE,show.legend=TRUE,color='black',method = 'lm',na.rm=TRUE) +
#   #geom_text(aes(x=Year,y=`% nest success`,colour = factor(Source),label=Label),check_overlap = TRUE,na.rm=TRUE) +
#   facet_grid(rows='RiverSegment') +
#   theme_minimal() 
  

```

# Discussion

[To be written]

# References

Kozie K. 1996. [Bald Eagle Productivity for Wrangell-St Elias National Park and Preserve 1993-1996](https://irma.nps.gov/DataStore/Reference/Profile/2238328). National Park Service, Wrangell - St. Elias National Park. Unpublished Report. Copper Center, Alaska

Postupalsky, S. 1974. Raptor reproductive success: some problems with methods, criteria, and terminology, IN Raptor Research Report No. 2, Raptor Research Foundation, Vermilion, South Dakota, pp 21-31.

Putera, J. and S.D. Miller. 2018. [Protocol implementation plan for bald eagle monitoring in Wrangell-St. Elias National Park & Preserve, Central Alaska Network: Version 1.0](https://irma.nps.gov/DataStore/Reference/Profile/2256539). Natural Resource Report. NPS/CAKN/NRR—2018/1752. National Park Service. Fort Collins, Colorado.

```{r echo=FALSE,include=FALSE}
# This chunk copies the report to the J or O drive
CurrentDirectory = "C:/Development/WRST BAEA Reports R Markdown/"
From = paste(CurrentDirectory,"WRST-BAEA-Annual-Summary-Report.html",sep='')
To = paste(CurrentDirectory,SurveyYear," WRST BAEA Survey Summary.html",sep='') # ARCN

# Copy the file to the Network reports directory
file.copy(From, To, overwrite = TRUE)

```

## R Markdown Report Template

This template was generated automatically from an R Markdown script available at <https://github.com/NPS-ARCN-CAKN/WRST-BAEA-Annual-Report-R-Markdown>. The information in the report was generated from bald eagle survey data in the NPS WRST_Caribou SQL Server database (inpyugamsvm01\nuna:WRST_Caribou).

