---
title: ""
format:
  html:
    toc: true
  typst:
    toc: false
    number-sections: false
    columns: 1
---

```{r setup, include=FALSE,echo=FALSE,warning=FALSE}

knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(odbc)
library(sf)

# Set the SurveyName for the report here
SurveyYear = 2025

# Database connection
Connection = dbConnect(odbc(),Driver = "Sql Server",Server = "INP2300IRMADB01.nps.doi.net\\NTWK", Database = "CAKN_BaldEagles")

CurrentDate = format(Sys.Date(), "%A, %B %d, %Y")
TableCounter = 1
FigureCounter = 1
```

# DRAFT Quality Control Tests Results: Bald Eagle Survey, `r SurveyYear`

## *National Park Service, Alaska Region*

Scott D. Miller\
Information Technology Specialist\
U.S. Department of the Interior\
National Park Service, Arctic Inventory and Monitoring Network\
240 W. 5th Ave\
Anchorage, AK 99501\


`r CurrentDate`

# Overview

The National Park Service monitors bald eagle productivity and abundance in the Copper River valley, Alaska, via aerial surveys. Data collected during these surveys are stored in a master database and undergo routine automated quality control tests to maximize data quality with results presented in this report.

## Spatial Data

A map of the survey area and muskox groups for which spatial coordinates were recorded appears in Figure `r FigureCounter`. Symbol size is representative of group size.

```{r,echo=FALSE,warning=FALSE,results='hide',include=FALSE}

# Get the observations having valid Latitude/Longitudes
Sql = paste0("SELECT  [Year]
      ,[TerritoryID]
      ,[NestNumber]
      ,[Nest_ID]
      ,[RiverSegment]
      ,[TreeSpecies]
      ,[Date_O]
      ,[Nest_O]
      ,[OccupancyStatus]
      ,[Date_P]
      ,[StatusCode_P]
      ,[ProductivityStatus]
      ,[NestIsNew]
      ,[NestIsGone]
      ,[NestIsGoneConfirmation]
      ,[Latitude]
      ,[Longitude]
  FROM [CAKN_BaldEagles].[dbo].[Dataset_Productivity]
  WHERE Year = ",SurveyYear)
obsdf = dbGetQuery(Connection,Sql)
obs = st_as_sf(obsdf %>% filter(!is.na(Latitude) & Latitude > 0),coords=c("Longitude","Latitude"),crs=4326)


# Get the parks and alaska outlines
Parks = st_read(r'(C:\Work\GIS Common Layers\AK_Parks.geojson)')
Parks = st_set_crs(Parks,4326)
AK = st_read(r'(C:\Work\GIS Common Layers\Alaska.geojson)')
AK = st_set_crs(AK,4326)

```

## Observations Map

```{r,echo=FALSE}
if(nrow(obs) > 0){
  # Get bounding box of the Plots layer
  bbox <- st_bbox(obs)
  buffer = 0.1
  ggplot() +
      
    # Alaska polygon
    geom_sf(data = AK, fill = "white", color = "gray80", alpha = 0.5) +
    
    # Parks polygons
    geom_sf(data = Parks %>% filter(PARK == 'bela' | PARK == 'cakr'), fill = "palegreen4", color = NA, alpha = 0.2) +

    # Park labels
    annotate("text", x = -164.2, y = 67.6, label = "Cape Krusenstern\nNational Monument",size = 3, color = "black", angle = 0,fontface='bold') +
    annotate("text", x = -164.5, y = 65.8, label = "Bering Land Bridge\nNational Preserve",size = 3, color = "black", angle = 0,fontface='bold') +
    
    # Observations
    geom_sf(data = obs,aes(color=StatusCode_P)) +
    
    # Bounding box restriction
    coord_sf(
      xlim = c(bbox["xmin"] - buffer, bbox["xmax"] + buffer),
      ylim = c(bbox["ymin"] - buffer, bbox["ymax"] + buffer),
      expand = FALSE
    ) +
    
    # Theming
    theme_minimal() +
    theme(
      panel.background = element_rect(fill = "lightsteelblue1"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    ) +
    labs(x='Longitude',y='Latitude',title='title')
}else{
  print("No spatial coordinates available for this survey")
}

```

## Bounding Box (Including NULLs)

```{r,echo=FALSE}
knitr::kable(t(obsdf %>% summarize(
    North = max(Latitude),
    South = min(Latitude),
    West = min(Longitude),
    East = max(Longitude),
    n=n()
  )
))

```
## Bounding Box (Excluding NULLs)

```{r,echo=FALSE}
knitr::kable(t(obsdf %>% summarize(
    `North` = max(Latitude,na.rm=TRUE),
    `South` = min(Latitude,na.rm=TRUE),
    `West` = min(Longitude,na.rm=TRUE),
    `East` = max(Longitude,na.rm=TRUE),
    n=n()
  )
))

```

# QC Test Results

```{r,echo=FALSE,results='asis'}
Tests = dbGetQuery(Connection,"SELECT[Table],[TableDescription],[SelectQuery] FROM [DatabaseTableDescriptions] WHERE Left([Table],3)='QC_' ORDER BY [Table]
")
i = 1
for(Table in Tests$Table){
  cat("## ",Table)
  cat("\n\n")
  TestDescription = Tests[i,"TableDescription"]
  cat("Description: ",TestDescription,"\n\n")
  Sql = paste0("SELECT Count(*) as n FROM ",Table," WHERE Year = ",SurveyYear)
  cat(Sql,"\n\n")
  Result = dbGetQuery(Connection,Sql)
  if(Result$n > 0){
    cat("\U1F534","<span style=\"color:red;\"> **FAILED**</span>\n\n")
  }else{
    cat("\U2705","Passed\n\n")
  }
  cat("\n\n")
  i=i+1
}

```

```{r,echo=FALSE,warning=FALSE,include=FALSE}
# This section copies the output file to a new html file titled with the survey name
# R doesn't seem to run this chunk during a knit operation so you'll have to set the SurveyName manually (way above) and then re-run this chunk to make sure the file gets copied correctly.
if(1==0){
  
  # Rename PDF
  dst <- paste0(SurveyName," Quality Control Report.pdf")
  src =  r'(C:\Development\ARCN Muskox R\ARCN Muskox Survey QC.pdf)'
  cat("Source file exists: ",file.exists(src),"\n")
  cat("Copy ",src," to ",dst,"\n")
  file.copy(src, dst, overwrite = TRUE)
  
  # Rename HTML
  dst <- paste0(SurveyName," Quality Control Report.html")
  src =  r'(C:\Development\ARCN Muskox R\ARCN Muskox Survey QC.html)'
  cat("Source file exists: ",file.exists(src),"\n")
  cat("Copy ",src," to ",dst,"\n")
  file.copy(src, dst, overwrite = TRUE)
  
}


```
