---
title: ""
format:
  typst:
    toc: false
    number-sections: false
    columns: 1
  html:
    toc: true

---

```{r setup, include=FALSE,echo=FALSE,warning=FALSE}

knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(odbc)
library(sf)

# Set the survey year for the report here
SurveyYear = 2025

# Database connection
Connection = dbConnect(odbc(),Driver = "Sql Server",Server = "INP2300IRMADB01.nps.doi.net\\NTWK", Database = "CAKN_BaldEagles")

CurrentDate = format(Sys.Date(), "%B %d, %Y")
TableCounter = 1
FigureCounter = 1

```

# Quality Control Tests Results: Bald Eagle Survey, `r SurveyYear`   
## *National Park Service*   
### Central Alaska Inventory and Monitoring Network      
   
   
   
   
Scott D. Miller   
Information Technology Specialist   
U.S. Department of the Interior   
240 W. 5th Ave   
Anchorage, AK 99501   


`r CurrentDate`

# Overview

The National Park Service monitors bald eagle productivity and abundance in the Copper River valley, Alaska, via aerial surveys. Information about the monitoring can be found in the [NPS Data Store](https://irma.nps.gov/DataStore/Reference/Profile/2216071). Data collected during occupance and productivity surveys are stored in a master database and undergo routine automated quality control tests to maximize data quality. Results of these tests are presented in this report.

## Spatial Data

A map of the survey area and muskox groups for which spatial coordinates were recorded appears in Figure `r FigureCounter`. Symbol size is representative of group size.

```{r,echo=FALSE,warning=FALSE,results='hide',include=FALSE}
# Load spatial layers from either Data Store (slow, but reliable for anyone anywhere) or local directory
if(1==1){
  # Load spatial files locally
  ShpPath = "C:/Work/GIS Common Layers/"
  Parks = read_sf(paste(ShpPath,"GCS/AKParks_reprojected.shp",sep=""))
  AK  = read_sf(paste(ShpPath,"GCS/AK_GCS_poly.shp",sep=""))
  Rivers = read_sf(paste0(ShpPath,"Alaska_Rivers_1M.geojson.geojson"))
  Roads = read_sf(paste0(ShpPath,"DOT_RoadSystem_091318/","DOT_RoadSystem_091318.shp"))
}else{
  # Load spatial files from Data Store
  Parks = st_read('https://irma.nps.gov/DataStore/DownloadFile/704836')
  AK = st_read('https://irma.nps.gov/DataStore/DownloadFile/746052')
  Rivers = read_sf('https://irma.nps.gov/DataStore/DownloadFile/758201') 
}

```

```{r,echo=FALSE,warning=FALSE,results='hide',include=FALSE}

# Get the observations having valid Latitude/Longitudes
Sql = paste0("SELECT  [Year]
      ,[TerritoryID]
      ,[NestNumber]
      ,[Nest_ID]
      ,[RiverSegment]
      ,[TreeSpecies]
      ,[Date_O]
      ,[Nest_O]
      ,[OccupancyStatus]
      ,[Date_P]
      ,[StatusCode_P]
      ,[ProductivityStatus]
      ,Productivity -- The actual classification field for productivity
      ,[NestIsNew]
      ,[NestIsGone]
      ,[NestIsGoneConfirmation]
      ,[Latitude]
      ,[Longitude]
  FROM [CAKN_BaldEagles].[dbo].[Dataset_Productivity]
  WHERE Year = ",SurveyYear)
obsdf = dbGetQuery(Connection,Sql)
obs = st_as_sf(obsdf %>% filter(!is.na(Latitude) & Latitude > 0),coords=c("Longitude","Latitude"),crs=4326)
obs %>% distinct(Productivity)
```

## Observations Map

```{r,echo=FALSE}
if(nrow(obs) > 0){
  # Get bounding box of the Plots layer
  bbox <- st_bbox(obs)
  buffer = 0.1
  ggplot() +
      
    # Alaska polygon
    #geom_sf(data = AK, fill = "white", color = "gray80", alpha = 0.5) +
    geom_sf(data=Parks,fill='gray80') +
    geom_sf(data = Rivers, fill = "lightsteelblue", color = 'lightsteelblue',alpha=0.6) +
    geom_sf(data = Roads, fill = "gray50", color = 'gray50',alpha=0.4) +
    
    # Parks polygons
    geom_sf(data = Parks %>% filter(PARK == 'bela' | PARK == 'cakr'), fill = "palegreen4", color = NA, alpha = 0.2) +

    # Park labels
    annotate("text", x = -164.2, y = 67.6, label = "Cape Krusenstern\nNational Monument",size = 3, color = "black", angle = 0,fontface='bold') +
    annotate("text", x = -164.5, y = 65.8, label = "Bering Land Bridge\nNational Preserve",size = 3, color = "black", angle = 0,fontface='bold') +
    
    # Observations
    geom_sf(data = obs,aes(color=Productivity,size=Productivity,shape=Productivity)) +
    scale_color_manual(values = c('Not initiated' = 'gray10','Successful' = 'green','Failed' = 'red','Productivity unknown' ='orange','Not found or not checked' = 'gray70')) +
    scale_shape_manual(values = c('Not initiated' = 1,'Successful' = 16,'Failed' = 16,'Productivity unknown' = 10,'Not found or not checked' = 10)) +
    scale_size_manual(values = c('Not initiated' = 1,'Successful' = 3,'Failed' = 3,'Productivity unknown' = 1,'Not found or not checked' = 1)) +

    # Bounding box restriction
    coord_sf(
      xlim = c(bbox["xmin"] - 1, bbox["xmax"] + 1),
      ylim = c(bbox["ymin"] - buffer, bbox["ymax"] + buffer),
      expand = FALSE
    ) +
    
    # Theming
    theme_minimal() +
    labs(x='Longitude',y='Latitude',title='Bald Eagle Observations')  +
    theme(axis.text.x = element_text(angle = 90,hjust = 1,vjust = 0.5))
  
}else{
  print("No spatial coordinates available for this survey")
}

```


## Bounding Box

```{r,echo=FALSE}
knitr::kable(t(obsdf %>% summarize(
    `North` = max(Latitude,na.rm=TRUE),
    `South` = min(Latitude,na.rm=TRUE),
    `West` = min(Longitude,na.rm=TRUE),
    `East` = max(Longitude,na.rm=TRUE),
    n=n()
  )
))

```

# Quality Control Test Results

# Nests

```{r,echo=FALSE,results='asis'}
Tests = dbGetQuery(Connection,"SELECT[Table],QC_Severity,[TableDescription],[SelectQuery] 
   FROM [DatabaseTableDescriptions] 
   WHERE Left([Table],8)='QC_Nests' 
   ORDER BY [Table]")
i = 1
for(Table in Tests$Table){
  tryCatch(
    {
        # Run the QC Test
        cat("## ",Table)
        cat("\n\n")
        TestDescription = Tests[i,"TableDescription"]
        cat("Description: ",TestDescription,"\n\n")
        Sql = paste0("SELECT Count(*) as n FROM ",Table)
        cat("SELECT * FROM \\[",Table,"\\] WHERE Year = ",SurveyYear,sep="")
        cat("\n\n")
        Result = dbGetQuery(Connection,Sql)
        
        # Present results
        if(Result$n > 0){
          if(Result$QC_Severity == 'Failure'){
            cat("\U1F621","<span style=\"color:red;\"> **FAILED**</span>\n\n")
          } else {
            cat("\U1F61E","<span style=\"color:orange;\"> **WARNING**</span>\n\n")
          }
          
        }else{
          cat("\U2705","<span style=\"color:green;\"> **Passed**</span>\n\n")
        }
        
        cat("\n\n")
        i=i+1
        
    },
    warning = function(w) {
      cat("\U1F7E1 <span style=\"color:yellow;\">Warning</span>: This quality control procedure does not apply to an annual survey.") #, conditionMessage(w)
      cat("\n\n")
    },
    error = function(e) {
      cat("\U1F7E1 <span style=\"color:yellow;\">Warning</span>: There was an error or this quality control procedure does not apply to an annual survey.") #, conditionMessage(e)
      cat("\n\n")
    },
    finally = {
      #message("Finished attempt.")
    })

}

```

# `r SurveyYear` Survey Data

```{r,echo=FALSE,results='asis'}
Tests = dbGetQuery(Connection,"SELECT[Table],QC_Severity,[TableDescription],[SelectQuery] 
   FROM [DatabaseTableDescriptions] 
   WHERE Left([Table],13)='QC_SurveyData' 
   ORDER BY [Table]")
i = 1
for(Table in Tests$Table){
  tryCatch(
    {
        # Run the QC Test
        cat("## ",Table)
        cat("\n\n")
        TestDescription = Tests[i,"TableDescription"]
        cat("Description: ",TestDescription,"\n\n")
        Sql = paste0("SELECT Count(*) as n FROM ",Table," WHERE Year = ",SurveyYear)
        cat("SELECT * FROM \\[",Table,"\\] WHERE Year = ",SurveyYear,sep="")
        cat("\n\n")
        Result = dbGetQuery(Connection,Sql)
        
        # Present results
        if(Result$n > 0){
          if(Result$QC_Severity == 'Failure'){
            cat("\U1F621","<span style=\"color:red;\"> **FAILED**</span>\n\n")
          } else {
            cat("\U1F61E","<span style=\"color:orange;\"> **WARNING**</span>\n\n")
          }
          
        }else{
          cat("\U2705","<span style=\"color:green;\"> **Passed**</span>\n\n")
        }
        
        cat("\n\n")
        i=i+1
        
    },
    warning = function(w) {
      cat("\U1F7E1 <span style=\"color:yellow;\">Warning</span>: This quality control procedure does not apply to an annual survey.") #, conditionMessage(w)
      cat("\n\n")
    },
    error = function(e) {
      cat("\U1F7E1 <span style=\"color:yellow;\">Warning</span>: There was an error or this quality control procedure does not apply to an annual survey.") #, conditionMessage(e)
      cat("\n\n")
    },
    finally = {
      #message("Finished attempt.")
    })

}

```

```{r,echo=FALSE,warning=FALSE,include=FALSE}

if(1==0){
  # Rename PDF
  src =  paste0("WRST BAEA Annual Data Quality Report.pdf")
  file.exists(src)
  dst <- paste0("BE15 ",SurveyYear," Bald Eagle Survey Data Quality Report.pdf")
  file.exists(dst)
  file.copy(src, dst, overwrite = TRUE)
}


```
