---
title: ""
format:
  typst:
    toc: false
    number-sections: false
    columns: 1
  html: default
---

# Certified Dataset: Long-term Monitoring of Bald Eagle Abundance and Distribution in Wrangell - St. Elias National Park and Preserve

## National Park Service  
## Wrangell - St. Elias National Park and Preserve
## Central Alaska Inventory and Monitoring Network

Scott D. Miller^1^ and Kyle Cutting^2^

^1^ U.S. Department of the Interior\
Enterprise Data Directorate\
240 W. 5th Ave Anchorage\
Alaska 99501

^2^ National Park Service\
Wrangell St. Elias National Park and Preserve\
8 Richardson Hwy Mile 106\
Copper Center, AK 99573

`r format(Sys.Date(), "%B %d, %Y")`

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) # Prevent overwhelming messages from reaching user

# Load libraries
library(tidyverse)
library(odbc)
library(sf)

# Make kable show blanks for NAs
options(knitr.kable.NA = '')

```

```{r echo=FALSE,label="SET UP REPORT PARAMETERS HERE"}

# Build a database connection
#ConnectionString = 'Driver={SQL Server};Server=inpyugamsvm01\\nuna;Database=CAKN_BaldEagles;Trusted_connection=yes'
Connection = dbConnect(odbc(),Driver = "Sql Server",Server = "INP2300IRMADB01.nps.doi.net\\NTWK", Database = "CAKN_BaldEagles")

# Function to convert a date string to a long date
GetLongDate <- function(Date) {
  return(format(as.Date(Date), format="%B %d, %Y "))
}

```

```{r echo=FALSE,include=FALSE}
# Generate the Nests deliverable file
Sql = paste("SELECT [Nest_ID]
      ,[TerritoryID]
      ,[NestNumber]
      ,[IsActive]
      ,[DiscoverDate]
      ,[FirstSightingDate]
      ,[LastSightingDate]
      ,[DateDiscoveredGone]
      ,[Fate]
      ,[Duration (years)]
      ,[RiverSegment]
      ,[TreeSpecies]
      ,[Species]
      ,[TreeIsDead]
      ,[NumberOfTimesNotFoundInTheLast2Surveys]
      ,[OccupancyObservations]
      ,[NestInitiations]
      ,[ProductivityObservations]
      ,[NumberOfTimesSuccessful]
      ,[PctSuccessful]
      ,[AvgYoungPerSuccessfulNest]
      ,[EarliestOccupancyRecord]
      ,[LatestOccupancyRecord]
      ,[EarliestProductivityRecord]
      ,[LatestProductivityRecord]
      ,[Latitude]
      ,[Longitude]
      --,[Location]
      ,[NestSub]
      ,[Comments]
      ,[WKT]
      --,[PointFeature]
  FROM [CAKN_BaldEagles].[dbo].[Dataset_Nests]",sep="")
Nests = dbGetQuery(Connection,Sql)
CertifiedDataDirectory = r'(C:\Work\VitalSigns\CAKN Bald Eagles\Data\Certified Datasets\)'
NestsFilename = paste0("WRST BAEA Certified Nests Dataset Ver. ",Sys.Date(),".csv")
NestsFile = paste0(CertifiedDataDirectory,NestsFilename)
if(1==0){
  write.csv(Nests,NestsFile,quote=TRUE,na='',row.names = FALSE)
}


```

```{r echo=FALSE,include=FALSE}
# Generate the surveys deliverable file
Sql = paste("SELECT [Year]
      ,[TerritoryID]
      ,[NestNumber]
      ,[Nest_ID]
      ,[RiverSegment]
      ,[TreeSpecies]
      ,[Date_O]
      ,[Nest_O]
      ,[OccupancyStatus]
      ,[Date_P]
      ,[StatusCode_P]
      ,[ProductivityStatus]
      ,[Productivity]
      ,[NestIsNew]
      ,[NestIsGone]
      ,[NestIsGoneConfirmation]
      ,[NumAds_O]
      ,[AdBehav1_O]
      ,[AdBehav2_O]
      ,[NumAds_P]
      ,[AdBehav1_P]
      ,[AdBehav2_P]
      ,[NumEggs_O]
      ,[NumYng]
      ,[NumFlg]
      ,[Chick_Age]
      ,[NestCondition_O]
      ,[Comments_O]
      ,[Comments_P]
      ,[DataProcessingComment]
      ,[SourceFile]
      ,[Latitude]
      ,[Longitude]
      ,[CertificationLevel]
      ,[CertificationDate]
      ,[CertifiedBy]
  FROM [CAKN_BaldEagles].[dbo].[Dataset_Productivity]
  WHERE CertificationLevel <> 'Raw'
  ORDER BY Year,TerritoryID,NestNumber",sep="")
SurveyData = dbGetQuery(Connection,Sql)
SurveyDataFilename = paste0("WRST BAEA Certified Productivity Dataset Ver. ",Sys.Date(),".csv")
SurveyDataFile = paste0(CertifiedDataDirectory,SurveyDataFilename)
if(1==0){
  write.csv(SurveyData,SurveyDataFile,quote=TRUE,na='',row.names = FALSE)
}

```

```{r echo=FALSE,include=FALSE}
# This query gets the survey details from the GSPE_Surveys database table
Sql = paste("SELECT TOP (1000) [Year]
      ,[MinOccupancySurveyDate] as [Occ. survey start]
      ,[MaxOccupancySurveyDate] as [Occ. survey start]
      ,[NumberOfOccupancySurveyDays] as [Days]
      ,[MinProductivitySurveyDate] as [Prod. survey start]
      ,[MaxProductivitySurveyDate] as [Prod. survey start]
      ,[NumberOfProductivitySurveyDays] as [Days]
      ,[NumberOfRecords] as Records
  FROM [CAKN_BaldEagles].[dbo].[Summary_SurveyDatesByYear]",sep="")
DF = dbGetQuery(Connection,Sql)

FirstSurveyYear = min(DF$Year)
LatestSurveyYear = max(DF$Year)

```

# Abstract

This dataset consists of certified nesting and productivity data from the National Park Service's long-term Bald Eagle monitoring program at Wrangell-St. Elias National Park and Preserve (WRST), Alaska, from `r FirstSurveyYear` to `r LatestSurveyYear`.

The Copper River drainage in WRST hosts a high density of nesting Bald Eagles *(Haliaeetus leucocephalus)*. Bald Eagles in WRST are a high profile species that are dependent upon many resources, including salmon, along the Copper River and nest at the northern edge of the their range. Bald Eagles are top-trophic level predators and they often respond quickly to changes in their environment by changing their breeding activities.

The objectives of the monitoring efforts are to 1) document the annual changes in occupancy of bald eagles, 2) document the annual changes in productivity of bald eagle nests, and document any changes in distribution of bald eagle nests.

# Methods

The bald eagle survey is part of long-term monitoring since 2004 (Putera and Miller, 2018). Five areas covering 378 miles of riverine habitat along the Copper, Chitina, and Bremner Rivers were surveyed (Figure 1). Two aerial surveys following the river corridor are conducted - one to document territorial occupancy, and the other to document productivity. Surveys are flown using a small airplane flying at low level (\<100m AGL) and low speed (\<100 kph), with one experienced observer.

We collect the following data on each nest observed: number of adults, behavior of adults, number of eggs and/or nestlings, tree species, status, and nest location. All nests still intact from the previous year’s surveys are rechecked for occupancy. During the productivity (chick) survey, all previously occupied nests from the first survey are rechecked . We define occupancy as a bird in incubation posture or two adult birds at a nest or in the nest tree. For the productivity survey, we recheck all occupied nests from the first survey to see if chicks are present.

# Study Area

Five survey areas were defined along major riverways, and are described in Kozie (1996, Figure 1):

1.) The Upper Copper River (UCR, 80km) extends from Copper and Tanada lakes to the confluence of the Chistochina River. White spruce (Picea glauca) is the predominant tree species, with balsam poplar (Populus balsamifera) common on the lower stretches. The Middle Copper River (MCR, 179km) extends from the Chistochina River south to the confluence of the Chitina River.

2.) The Middle Copper River (MCR) contains a mixture of both white spruce and balsam poplar, with poplar being slightly more common. The Copper River begins to braid along the MCR.

2.) The Lower Copper River (LCR) section (188km) extends from the Chitina River south to Miles Lake. The LCR is dominated by cottonwood along the river. The vegetation also begins to exhibit coastal influences, adding several coastal species including black cottonwood (Populus trichocarpa), Sitka spruce (Picea sitchensis), and mountain hemlock (Tsuga mertensiana). The Copper River is wide and more braided than the upper sections, with steep canyons on either side. Nesting begins consistently later along the LCR and is probably influenced by the numerous glaciers in the area which delay snow melt and maintain the area in a winter-like state several weeks after green up has begun to occur to the north.

3.) The Bremner River (45km) is surveyed from its confluence with the Copper River to the forks by Threemile Canyon, and has similar characteristics as the LCR.

4.) The Chitina River (116km) is surveyed from the confluence of the Copper River to where the river becomes braided above the Tana River. White spruce dominate the trees, with cottonwood becoming scarce in the upper reaches.

A total of 609 km of riverways are surveyed annually, amounting to over 1200km of surveys.

```{r echo=FALSE}
# Load spatial layers from either Data Store (slow, but reliable for anyone anywhere) or local directory
if(1==1){
  # Load spatial files locally
  ShpPath = "C:/Work/GIS Common Layers/"
  Parks = read_sf(paste(ShpPath,"AK_Parks.geojson",sep=""))
  AK  = read_sf(paste(ShpPath,"Alaska.geojson",sep=""))
  Rivers = read_sf(paste0(ShpPath,"Alaska_Rivers_1M.geojson.geojson"))
  Roads = read_sf(paste0(ShpPath,"AK_Roads.geojson"))
  Lakes = read_sf(paste0(ShpPath,"AK_Lakes/mv_major_lake_py.shp"))
}else{
  # Load spatial files from Data Store
  Parks = st_read('https://irma.nps.gov/DataStore/DownloadFile/704836')
  AK = st_read('https://irma.nps.gov/DataStore/DownloadFile/746052')
  Rivers = read_sf('https://irma.nps.gov/DataStore/DownloadFile/758201') 
}

# Build a map of the survey area showing park boundaries, copper river and tributaries, etc.
ggplot() +
  
  # Blue rectangle to make the background of the map blue
  geom_rect(aes(xmin = -155,xmax = -140,ymin = 55,ymax = 70),fill = "steelblue3") +
  
  # Contextual layers
  geom_sf(data=AK,mapping=aes(), color='darkgray',fill='snow2') +
  geom_sf(data=Parks,mapping=aes(), color='yellow4',fill='yellow4',alpha=0.3) +
  geom_sf(data=Rivers %>% filter(TYPE != 'I'),mapping=aes(),color='steelblue3',fill='steelblue3') +
  geom_sf(data = Roads, fill = "gray50", color = 'gray50',alpha=0.4) +
  geom_sf(data=Lakes,mapping=aes(),color='steelblue3',fill='steelblue3') +
  
  # Set the map limits, zoom into the BAEA monitoring area
  coord_sf(xlim = c(-146.6, -142), ylim = c(60.5, 62.8), expand = FALSE) +
  xlab('Longitude') +
  ylab('Latitude') +

  # WRST text Label
  geom_text(aes(x = -144.25, y = 61.8, label = "Wrangell - St. Elias National\nPark and Preserve"), size = 4, color = "gray40", fontface = "bold") +
  
  # Upper Copper River label
  geom_text(aes(x = -144, y = 62.45, label = "Upper Copper River"), size = 4, color = "black") +
  
  # Middle Copper River label
  geom_text(aes(x = -145.2, y = 61.5, label = "Middle Copper River"), size = 4, color = "black") +
  
  # Lower Copper River label
  geom_text(aes(x = -145.2, y = 60.8, label = "Lower Copper River"), size = 4, color = "black") +
  
  # Chitina River label
  geom_text(aes(x = -143.8, y = 61.25, label = "Chitina River"), size = 4, color = "black") +
  
  # Bremner River label
  geom_text(aes(x = -143.8, y = 60.95, label = "Bremner River"), size = 4, color = "black") +
  
  # Minimal theme
  theme_minimal() +
  
  # Legend position
  theme(legend.position = "bottom")


```

Figure 1. Bald Eagles monitoring study area in Wrangell - St. Elias National Park and Preserve, Alaska.

# Data files

```{r, echo=FALSE}
CertifiedDataDirectory = r'(C:\Work\VitalSigns\CAKN Bald Eagles\Data\Certified Datasets)'
FilesList = as.data.frame(list.files(path = CertifiedDataDirectory, pattern = "\\.csv$", full.names = FALSE))
colnames(FilesList) = 'File'
FilesList$Description = 'Nests dataset'
FilesList[2,2] = 'Productivity dataset'

```

This data release consists of two files, `r FilesList[1,1]` and `r FilesList[2,1]` (Table 1).

Table 1. Files included in this release.

```{r, echo=FALSE}
knitr::kable(FilesList)

```

# Columns descriptions

## Nests dataset

Column descriptions and data types describing the Nests dataset appear in Tables 2 and 3, below.

Table 2. Column descriptions for the Nests dataset.

```{r , echo=FALSE}
Sql = "SELECT [Table],[Column],DataType,ColumnDescription as Description FROM Dataset_Certified_Metadata ORDER BY [Table]"
MD = dbGetQuery(Connection,Sql)
knitr::kable(MD %>% filter(Table == 'Dataset_Nests') %>% select(-Table,-DataType))

```

Table 3. Column data types for the Nests dataset.

```{r , echo=FALSE}
knitr::kable(MD %>% filter(Table == 'Dataset_Nests') %>% select(-Table,-Description))

```
## Productivity dataset

Column descriptions and data types describing the productivity dataset appear in Tables 4 and 5, below.

Table 4. Column descriptions for the productivity dataset.

```{r , echo=FALSE}
Sql = "SELECT [Table],[Column],DataType,ColumnDescription as Description FROM Dataset_Certified_Metadata ORDER BY [Table]"
MD = dbGetQuery(Connection,Sql)
knitr::kable(MD %>% filter(Table == 'Dataset_Productivity') %>% select(-Table,-DataType))

```

Table 5. Column data types for the productivity dataset.

```{r , echo=FALSE}
knitr::kable(MD %>% filter(Table == 'Dataset_Productivity') %>% select(-Table,-Description))

```

# Data Summary

A summary of the survey dates and number of records appears in Table 6, below.



```{r , echo=FALSE}
Sql = "SELECT Year
, Records
--, [Available nests]
--, [Nests located]
, [Start of productivity surveys] as [Start]
, [End of productivity surveys] as [End]
, [Number of productivity survey days] as Days
FROM            Dashboard
ORDER BY Year"
Summary = dbGetQuery(Connection,Sql)

```

The Bald Eagle monitoring dataset consists of `r nrow(Summary)` Bald Eagle surveys (Table 6).

Table 6. Summary of survey dates and number of records in the Bald Eagles monitoring dataset.

```{r , echo=FALSE}
knitr::kable(Summary)
      
```

# Reference

Putera, J. and S.D. Miller. 2018. Protocol implementation plan for bald eagle monitoring in Wrangell-St. Elias National Park & Preserve, Central Alaska Network: Version 1.0. Natural Resource Report. NPS/CAKN/NRR—2018/1752. National Park Service. Fort Collins, Colorado https://irma.nps.gov/DataStore/Reference/Profile/2256539
