---
title: ""
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

```{=html}
<style>

body {
    font-family: "Times New Roman";
    font-size: 11.5pt;
    color: black;
}

.title {
  font-family: "Times New Roman";
  font-size: 16pt;
  font-weight: bold;
  color: black;
}

h1 {
    font-family: "Arial";
    font-size: 16pt;
    font-weight: bold;
    color: black;
}


h2 {
  font-family: "Arial";
  font-size: 11.5pt;
  font-weight: bold;
  color: black;
}

h3 {
  font-family: "Arial";
  font-size: 10.5pt;
  font-weight: bold;
  color: black;
}

h4 {
    font-family: "Times New Roman";
    font-size: 12pt;
    color: black;
}

p {
    font-family: "Times New Roman";
    font-size: 11.5pt;
    color: black;
}

.caption {
    font-family: "Arial";
    font-size: 10pt;
    color: black;
}

th {
    font-family: "Arial";
    font-size: 9pt;
    font-weight: bold;
    color: black;
}

tr,td,table {
    font-family: "Arial";
    font-size: 9pt;
    color: black;
}

.header{
  font-family: "Arial";
  font-size: 10pt;
  font-weight: bold;
  color: black;
}

</style>
```

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) # Prevent overwhelming messages from reaching user

options(download.file.method = "wininet")
#install.packages("sqldf")
# Load libraries
library(odbc)
library(knitr)
library(ggplot2)
library(sp)
library(sf)
library(sqldf)
library(tidyverse)


# Make kable show blanks for NAs
options(knitr.kable.NA = '')

# These parameters restrict the dimensions of plots to 6 metric inches so they don't blow out Word
FigWidth = 6
```

```{r echo=FALSE,label="SET UP REPORT PARAMETERS HERE"}

# Set up report parameters here
SurveyYear = 2024  # Set this to the desired report year. Don't set to 2016 because that survey didn't come off and the script will bomb

# Database connection
Connection = dbConnect(odbc(),Driver = "Sql Server",Server = "inpyugamsvm01\\nuna", Database = "CAKN_BaldEagles")

# Set the path to shapefiles used in the map, AK, Parks, Rivers, etc
ShpPath = "C:/Work/GIS Common Layers/"

```



```{r echo=FALSE,label="Functions"}

# Current date
CurrentDate = format(Sys.Date(), format='%B %d, %Y') # Get the current date

# Function to convert a date string to a long date
GetLongDate <- function(Date) {
  return(format(as.Date(Date), format="%B %d, %Y"))
}

GetQCSection = function(Sql){
  cat("## QC-X: QCTestDescription  \n",sep="")
  cat("\n",sep="")
  cat("DESCRIPTION  \n",sep="")
  cat("\n",sep="")
  cat("",sep="")
  cat("```{r echo=FALSE}  \n",sep="")
  cat("Sql = paste(\"",Sql,"\",sep=\"\")\n",sep="")
  cat("df = dbGetQuery(Connection,Sql)  \n",sep="")
  cat("if(nrow(df)==0){  \n",sep="")
    cat("    knitr::asis_output(\"*PASSED*. Zero records returned.  \")  \n",sep="")
  cat("} else{  \n",sep="")
    cat("    knitr::kable(df)  \n",sep="")
  cat("}  \n",sep="")
  cat("```  \n",sep="")
  cat("`r Sql`\n\n",sep="") 
  cat("`r nrow(df)` rows  \n\n",sep="")
}
#Test
# GetQCSection("select x from y")

# These counters are used to number the tables and figures. Each gets incremented after use.
TableCounter = 1
FigureCounter = 1

```


# Post-survey Data Quality Report For The Bald Eagles Monitoring Program in Wrangell-St. Elias National Park and Preserve, Alaska

Scott D. Miller  
National Park Service, Arctic Inventory and Monitoring Network  
240 W. 5th Ave  
Anchorage, Alaska 99501   

`r CurrentDate`

U.S. Department of the Interior\
National Park Service, Alaska Region\
Anchorage, Alaska\

# Introduction

In the realm of biological sciences, the integrity of data is crucial for advancing our understanding of complex biological processes and ecosystems. Ensuring high data quality is fundamental for producing reliable results, validating hypotheses, and making informed decisions for the management of natural areas, enhancing conservation efforts, and communicating results to the public. Data quality control is a critical step in the scientific life cycle for long-term monitoring of ecosystems, ensuring the accuracy, consistency, and completeness of data collected. This report documents the quality control procedures executed on the field data collected during the `r SurveyYear` bald eagles monitoring survey in Wrangell-St. Elias National Park and Preserve in Alaska.

Each section of this report details a quality control check performed on the master bald eagles monitoring database, along with any findings. Any identified quality control defects should be fixed or documented. The report is broken out into two major sections related to 1) information about eagle nests and 2) survey results.

# Quality Control Tests: Nests

## QC-N01: Active nests with DateDiscoveredGone is not NULL

All active nests should have DateDiscoveredGone attributes NULL

```{r echo=FALSE}
Sql = paste("SELECT Nest_ID, Date_O, Nest_O, Date_P, StatusCode_P, Comments_O, Comments_P, IsActive, DateDiscoveredGone
FROM QC_ActiveNestsWithDateDiscoveredGoneIsNotNULL
WHERE Year(Date_O) = ",SurveyYear,sep='')
df = dbGetQuery(Connection,Sql)
if(nrow(df)==0){
  knitr::asis_output("*PASSED*. Zero records returned.  \n")
} else{
  knitr::kable(df)
}


```


`r Sql` 

`r nrow(df)` rows

## QC-N02: Nests With No Productivity Observation

Occasionally observers are unable to re-survey nests for the productivity portion of the survey due to bad weather, smoke, logistical, or other problems. These records should have their StatusCode_P attribute set to 'NS' (Not surveyed). The table below shows records with StatusCode_P = 'NS'

```{r echo=FALSE}
Sql = paste("SELECT Nest_ID, RiverSegment, Date_O, OccupancyStatus, ProductivityStatus,Comments_O,Comments_p
FROM QC_NestsNotProductivitySurveyed
WHERE (NOT (Date_O IS NULL)) AND (Date_P IS NULL)
And Year(Date_O)=",SurveyYear," and StatusCode_P='NS'
ORDER BY Date_O,Nest_ID",sep='')
df = dbGetQuery(Connection,Sql)
if(nrow(df)==0){
  knitr::asis_output("*PASSED*. Zero records returned.  \n")
} else{
  knitr::kable(df)
}
```

`r Sql`

`r nrow(df)` rows





## QC-N03: Active nests that possibly should be inactive

The following nests, if any, should possibly be inactivated.

```{r echo=FALSE}
Sql = paste("SELECT        Nest_ID, IsActive, DiscoverDate, LastSightingDate, NumberOfTimesNotFoundInTheLast2Surveys, Comments
--, HelperQuery, UpdateQuery
FROM            QC_NestsThatProbablyShouldBeInactivated",sep='')
df = dbGetQuery(Connection,Sql)
if(nrow(df)==0){
  knitr::asis_output("*PASSED*. Zero records returned.  \n")
} else{
  knitr::kable(df)
}

```

`r Sql`

`r nrow(df)` rows


## QC-N04: Minimum/Maximum Nest Spatial Coordinates

This section shows the minimum and maximum spatial coordinates of nests. The objective is to debug nests with coordinates equal to zero, or that are NULL. 

```{r echo=FALSE}  
Sql = paste("SELECT TOP (1000) [Minimum latitude]
      ,[Maximum latitude]
      ,[Minimum longitude]
      ,[Maximum longitude]
      ,[Minimum non-zero latitude]
      ,[Maximum non-zero latitude]
      ,[Minimum non-zero longitude]
      ,[Maximum non-zero longitude]
  FROM [CAKN_BaldEagles].[dbo].[QC_Nests_MinMaxBoundingBoxCheck]",sep="")
df = dbGetQuery(Connection,Sql)  
knitr::kable(t(df))  

```  


## QC-N05: Nests missing spatial coordinates  

The following nests are missing spatial coordinates.  

```{r echo=FALSE}  
Sql = paste("SELECT [Nest_ID],[Lon-NAD83],[Lat-NAD83],[Comments]  FROM [CAKN_BaldEagles].[dbo].[QC_Nests_MissingCoordinates] ORDER BY Nest_ID",sep="")
df = dbGetQuery(Connection,Sql)  
if(nrow(df)==0){  
    knitr::asis_output("*PASSED*. Zero records returned.  ")  
} else{  
    knitr::kable(df)  
}  
```  

## QC-N06: Active nests with DateDiscoveredGone is not NULL

Active nests should have DateDiscoveredGone = NULL  

```{r echo=FALSE}  
Sql = paste("SELECT [Nest_ID],[IsActive],[DateDiscoveredGone],[Comments]
-- ,NestDebugger 
FROM [CAKN_BaldEagles].[dbo].[QC_Nests_NestIsActiveButDateDiscoveredGoneIsNotNull] ORDER BY Nest_ID",sep="")
df = dbGetQuery(Connection,Sql)  
if(nrow(df)==0){  
    knitr::asis_output("*PASSED*. Zero records returned.  ")  
} else{  
    knitr::kable(df)  
} 

```  

## QC-N07: Nests with no observations

The following nests have no observations. They may have been lost the same survey year they were discovered.

```{r echo=FALSE}  
Sql = paste("SELECT Nest_ID, DiscoverDate, TreeIsDead, DateDiscoveredGone, Comments FROM QC_Nests_NestsWithNoObservations",sep="")
df = dbGetQuery(Connection,Sql)  
if(nrow(df)==0){  
    knitr::asis_output("*PASSED*. Zero records returned.  ")  
} else{  
    knitr::kable(df)  
}  
```  
`r Sql`

`r nrow(df)` rows  



# Quality Control Tests: Survey Observations

## QC-S01: Active nests that were not search for during the latest survey

The results below active nests, if any, that were not searched for during the latest survey.

```{r echo=FALSE}
Sql = paste("SELECT [Nest_ID]
      ,[RiverSegment]
      ,[LastSightingDate]
      ,[IsActive]
      ,[NumberOfTimesNotFoundInTheLast2Surveys]
      ,[Comments]
 --     ,[UPDATEQuery]
--      ,[HelperQuery]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_ActiveNestsNotSearchedForDuringLastSurvey]",sep='')
df = dbGetQuery(Connection,Sql)
if(nrow(df)==0){
  knitr::asis_output("*PASSED*. Zero records returned.  \n")
} else{
  knitr::kable(df)
}

```

`r Sql`

`r nrow(df)` rows


## QC-S02: Inactive nests that possibly should be active

The following nests, if any, should possibly be inactivated.

```{r echo=FALSE}
Sql = paste("SELECT  [Nest_ID]
      ,[IsActive]
      ,[NumberOfTimesNotFoundInTheLast2Surveys]
      ,[LastSightingDate]
      ,[Comments]
--      ,[SELECTQuery]
--     ,[NestsQuery]
--      ,[UPDATEQuery]
  FROM [CAKN_BaldEagles].[dbo].[QC_Nests_InActiveNestsThatPossiblyShouldBeActive]
ORDER BY LastSightingDate",sep='')
df = dbGetQuery(Connection,Sql)
if(nrow(df)==0){
  knitr::asis_output("*PASSED*. Zero records returned.  \n")
} else{
  knitr::kable(df)
}
```

`r Sql`

`r nrow(df)` rows



## QC-S03: Possibly mistaken nest initiation decisions

Initiated nests should have at least two adults present during the occupancy survey with incubating behaviors observed. The following observations may have been mistakenly ruled as initiated nests.

```{r echo=FALSE}
Sql = paste("SELECT [Year]
      ,[Nest_ID]
      ,[Date_O]
      ,[Nest_O]
      ,[NumEggs_O]
      ,[NumAds_O]
      ,[AdBehav1_O]
      ,[AdBehav2_O]
      ,[NumYng]
      ,[StatusCode_P]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_PossiblyBadNestInitiations]
  WHERE Year=",SurveyYear," ORDER BY Nest_ID",sep='')
df = dbGetQuery(Connection,Sql)
if(nrow(df)==0){
  knitr::asis_output("*PASSED*. Zero records returned.  \n")
} else{
  knitr::kable(df)
}

```

`r Sql`

`r nrow(df)` rows


## QC-S04: Nests marked successful that did not produce any young

Nests marked successful that did not produce any young.

```{r echo=FALSE}
Sql = paste("SELECT [Nest_ID]
      ,[NumYng]
      ,[StatusCode_P]
      --,Year
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_SuccessfulNestsWithNoYoung] WHERE Year=",SurveyYear," ORDER BY Nest_ID",sep='')
df = dbGetQuery(Connection,Sql)
if(nrow(df)==0){
  knitr::asis_output("*PASSED*. Zero records returned.  \n")
} else{
  knitr::kable(df)
}

```

`r Sql`

`r nrow(df)` rows



## QC-S05: Chick Age Recorded But No Young Recorded

The following records, if any, have a Chick_Age attribute that is not NULL, but NumYng attribute is NULL or zero. This means a chick stage was recorded, but no chicks were observed.

```{r echo=FALSE}
Sql = paste("SELECT        Year, Nest_ID, Date_O, Nest_O, Date_P, StatusCode_P, NumYng, Chick_Age, Comments_O, Comments_P
FROM            QC_ChickAgeButNoYoung WHERE Year = ",SurveyYear,sep='')
df = dbGetQuery(Connection,Sql)
if(nrow(df)==0){
  knitr::asis_output("*PASSED*. Zero records returned.  \n")
} else{
  knitr::kable(df)
}

```

`r Sql`

`r nrow(df)` rows


## QC-S06: Successful Nests With No Recorded Young

The definition of nest success is observed young during the productivity survey. The following records, if any, should be checked.

```{r echo=FALSE}
Sql = paste("SELECT [Year]
      ,[Nest_ID]
      ,[NumYng]
      ,[StatusCode_P]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_SuccessfulNestsWithNoYoung]
  WHERE Year = ",SurveyYear,sep='')
df = dbGetQuery(Connection,Sql)
if(nrow(df)==0){
  knitr::asis_output("*PASSED*. Zero records returned.  \n")
} else{
  knitr::kable(df)
}

```

`r Sql`

`r nrow(df)` rows


## QC-S07: Possibly incorrect NestIsGoneConfirmation  

Each observation after the one where the nest is observed gone should have its NestIsGoneConfirmation attribute set to 1 in order to exclude it from analyses. This test returns nests that violate this premise. Use the NestDebugger stored procedure to troubleshoot the problem.

```{r echo=FALSE}  
Sql = paste("SELECT TOP (1000) [Nest_ID]
      ,[YearNestIsGone]
      --,[FixNestIsGoneConfirmation]
      ,[NestDebugger]
      --,[UpdateQuery]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_IncorrectNestIsGoneConfirmation]",sep="")
df = dbGetQuery(Connection,Sql)  
if(nrow(df)==0){  
    knitr::asis_output("*PASSED*. Zero records returned.  ")  
} else{  
    knitr::kable(df)  
}  

```  
`r Sql`

`r nrow(df)` rows  


## QC-S08: Initiated nests missing follow up

The following records are missing follow up productivity survey information (date, productivity code or productivity metrics).  

```{r echo=FALSE}  
Sql = paste("SELECT TOP (1000) [Year]
      ,[Nest_ID]
      ,[Date_O]
      ,[Date_P]
      ,[Nest_O]
      ,[StatusCode_P]
      ,[Comments_O]
      ,[Comments_P]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_InitiatedNestsMissingFollowUp]",sep="")
df = dbGetQuery(Connection,Sql)  
if(nrow(df)==0){  
    knitr::asis_output("*PASSED*. Zero records returned.  ")  
} else{  
    knitr::kable(df)  
}  
```  
`r Sql`

`r nrow(df)` rows  


## QC-S09: Multiple NestIsGone = 1  

Each Nest should only have one NestIsGone attribute set to 1 (true) to indicate the year the nest was observed gone. Subsequent observations to verify the nest is gone should have their NestIsGoneConfirmation attributes set to 1 so they can be filtered out of analyses since they hold no data. This view shows nest records having multiple records having NestIsGone = 1

EXEC NestDebugger will show pertinent Nest and SurveyData records for troubleshooting 

EXEC SetNestIsGone will output some Update queries that may help to fix the problems rapidly.

```{r echo=FALSE}  
Sql = paste("SELECT [Nest_ID]
      ,[NestIsGoneCount]
      ,[NestDebugger]
      ,[FixNestIsGone]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_MultipleNestIsGoneDates] ORDER BY Nest_ID",sep="")
df = dbGetQuery(Connection,Sql)  
if(nrow(df)==0){  
    knitr::asis_output("*PASSED*. Zero records returned.  ")  
} else{  
    knitr::kable(df)  
}  

```  
`r Sql`

`r nrow(df)` rows  


## QC-S10: Multiple SurveyData.NestIsNew = 1 records

Only the first nest observation should have its SurveyData.NestIsNew attribute set to 1.  

```{r echo=FALSE}  
Sql = paste("SELECT [Nest_ID]
      ,[NestIsNew]
      ,[n]
      ,[NestDebugger]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_MultipleNestIsNewValues]
ORDER BY Nest_ID
",sep="")
df = dbGetQuery(Connection,Sql)  
if(nrow(df)==0){  
    knitr::asis_output("*PASSED*. Zero records returned.  ")  
} else{  
    knitr::kable(df)  
}  


```  
`r Sql`

`r nrow(df)` rows  


## QC-S11: Multiple records with NestIsGone attribute set to 1

A nest can only be gone once. This test shows nests with multiple SurveyDate.NestIsGone attributes = 1.

```{r echo=FALSE}  
Sql = paste("SELECT [Nest_ID]
      ,[NestIsGoneCount]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_MultipleRecordsHavingNestIsGoneGT1]",sep="")
df = dbGetQuery(Connection,Sql)  
if(nrow(df)==0){  
    knitr::asis_output("*PASSED*. Zero records returned.  ")  
} else{  
    knitr::kable(df)  
}  
```  
`r Sql`

`r nrow(df)` rows  



## QC-S12: Possibly bad nest initiation assessments

Nest initiation is defined as one or more of the following:

- Two adults by the nest  
- Incubating posture in one adult  
- Visible egg(s) during the occupancy survey  

The records below violate these conditions.

```{r echo=FALSE}  
Sql = paste("SELECT  [Year]
      ,[Nest_ID]
      ,[Nest_O]
      ,[StatusCode_P]
      ,[AdBehav1_O]
      ,[AdBehav2_O]
      ,[NumEggs_O]
      ,[NumAds_O]
      ,[NumYng]
      --,[Date_O]
      --,[Date_P]
      ,[Comments_O]
      ,[Comments_P]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_PossiblyBadNestInitiations] ORDER BY Year,Nest_ID",sep="")
df = dbGetQuery(Connection,Sql)  
if(nrow(df)==0){  
    knitr::asis_output("*PASSED*. Zero records returned.  ")  
} else{  
    knitr::kable(df)  
}  
```  
`r Sql`

`r nrow(df)` rows



## QC-S13: Nests marked successful but did not produce young 

Successful nests must have produced young to be defined as successful. The following records violate this rule.  

```{r echo=FALSE}  
Sql = paste("SELECT [Year]
      ,[Nest_ID]
      ,[NumYng]
      ,[StatusCode_P]
  FROM [CAKN_BaldEagles].[dbo].[QC_SurveyData_SuccessfulNestsWithNoYoung]
ORDER BY Year,Nest_ID",sep="")
df = dbGetQuery(Connection,Sql)  
if(nrow(df)==0){  
    knitr::asis_output("*PASSED*. Zero records returned.  ")  
} else{  
    knitr::kable(df)  
}  
```  
`r Sql`

`r nrow(df)` rows  




## QC Appendix: List of comments

Comments can be useful in troubleshooting strange results

```{r echo=FALSE}

Sql=paste("SELECT Nest_ID,Comments_O as Comment,'Occupancy' as Survey  
FROM SurveyData 
WHERE Year = ",SurveyYear," And (NOT (Comments_O IS NULL))
UNION
SELECT Nest_ID,Comments_P,'Productivity' 
FROM SurveyData 
WHERE Year = ",SurveyYear," And (NOT (Comments_P IS NULL))
ORDER BY Nest_ID",sep="")
df = dbGetQuery(Connection,Sql)
if(nrow(df)==0){
  knitr::asis_output("*PASSED*. Zero records returned.  \n")
} else{
  knitr::kable(df)

}

```

## R Markdown Report Template

This template was generated automatically from an R Markdown script available at <https://github.com/NPS-ARCN-CAKN/WRST-BAEA-Annual-Report-R-Markdown>. The information in the report was generated from bald eagle survey data in the NPS WRST_Caribou SQL Server database (inpyugamsvm01\nuna:WRST_Caribou).

