---
title: ""
format:
  typst:
    toc: false
    number-sections: false
    columns: 1
  html: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) # Prevent overwhelming messages from reaching user
options(download.file.method = "wininet")
# Make kable show blanks for NAs
options(knitr.kable.NA = '')


# Load libraries
library(odbc)
library(knitr)
# library(ggplot2)
# library(sp)
library(sf)
# library(sqldf)
library(tidyverse)



# These parameters restrict the dimensions of plots to 6 metric inches so they don't blow out Word
FigWidth = 6
```

```{r echo=FALSE,label="SET UP REPORT PARAMETERS HERE"}

# Set up report parameters here
SurveyYear = 2025  # Set this to the desired report year. Don't set to 2016 because that survey didn't come off and the script will bomb

# ODBC connection to the BAEA monitoring SQL Server Database. 
# Contact data manager for server and database details.
# Connection to the bald eagles Sql Server database
Connection = dbConnect(odbc(),Driver = "Sql Server",Server = "INP2300IRMADB01.nps.doi.net\\NTWK", Database = "CAKN_BaldEagles")

# Set the path to shapefiles used in the map, AK, Parks, Rivers, etc
ShpPath = "C:/Work/GIS Common Layers/"

```

```{r echo=FALSE,label="Functions"}

# Current date
CurrentDate = format(Sys.Date(), format='%B %d, %Y') # Get the current date

# Function to return data frames from SQL database queries
GetDataFrameFromDatabase <- function(Sql) {
   return(dbGetQuery(Connection, Sql))
}

# Function to convert a date string to a long date
GetLongDate <- function(Date) {
  return(format(as.Date(Date), format="%B %d, %Y"))
}

# These counters are used to number the tables and figures. Each gets incremented after use.
TableCounter = 1
FigureCounter = 1

```

```{r echo=FALSE, label="Survey dates"}

# Summary_SurveyDatesByYearDataFrame: Survey Dates
Sql = paste("SELECT 
--[Year],
[MinOccupancySurveyDate] as [Occupancy survey start]
,[MaxOccupancySurveyDate] as [Occupancy survey end]
-- ,[NumberOfOccupancySurveyDays]
,[MinProductivitySurveyDate] as [Productivity survey start]
,[MaxProductivitySurveyDate] as [Productivity survey end]
-- ,[NumberOfProductivitySurveyDays]
FROM [CAKN_BaldEagles].[dbo].[Summary_SurveyDatesByYear] 
WHERE [Year] = ",SurveyYear,sep="")
Summary_SurveyDatesByYearDataFrame = GetDataFrameFromDatabase(Sql) # Get the data from the database into a data frame

# Build some survey date values to use later in the markdown text
OStart = Summary_SurveyDatesByYearDataFrame$`Occupancy survey start`
OEnd = Summary_SurveyDatesByYearDataFrame$`Occupancy survey end`
PStart = Summary_SurveyDatesByYearDataFrame$`Productivity survey start`
PEnd = Summary_SurveyDatesByYearDataFrame$`Productivity survey end`

# Summary_SuccessByYear: Success and productivity data by year
# Sql = paste("SELECT [Year]
# ,[Nests located]
# ,[New nests]
# ,[Nests lost]
# ,[Occupied]
# ,[Successful]
# ,[Productivity unknown]
# --,[Eggs]
# ,[Young fledged]
# ,[% nest occupied]
# ,[% nest success]
# ,[Young/occupied nest]
# ,[Young/successful nest]
# --,Source
# --,Left(Source,4) as Source
# FROM [Summary_SuccessByYear]
# WHERE Year <= ",SurveyYear," 
# AND Year NOT IN (2016) -- 2016 Interrupted by bad weather
# ORDER BY [Year] DESC",sep="")

Sql = paste("SELECT * FROM Dashboard WHERE Year <= ",SurveyYear," AND Year NOT IN (2016) ORDER BY [Year] DESC",sep="")

# Get the data from the database into a data frame
Summary_SuccessByYearDataFrame = GetDataFrameFromDatabase(Sql) 

# Length of record
MinYear = min(Summary_SuccessByYearDataFrame$Year, na.rm = TRUE) 
MaxYear = max(Summary_SuccessByYearDataFrame$Year, na.rm = TRUE)
LengthOfRecord = MaxYear - MinYear
FirstSurveyYear = min(Summary_SuccessByYearDataFrame$Year)
NumberOfSurveys = Summary_SuccessByYearDataFrame %>% summarize(n=n()) %>% pull(n) 

# We'll need just this year's data in a separated data frame from all the historical data.
# Filter on the summary data frame on the current year and then remove the Year column
Summary_SuccessByYearDataFrame_Current = Summary_SuccessByYearDataFrame %>% filter(Year == SurveyYear)


# Build some summary statistics to report in the markdown
NestsLocated = Summary_SuccessByYearDataFrame_Current$`Nests located`
Occupied = Summary_SuccessByYearDataFrame_Current$Occupied
Successful = Summary_SuccessByYearDataFrame_Current$Successful
YoungFledged = Summary_SuccessByYearDataFrame_Current$`Young`
NF = Summary_SuccessByYearDataFrame_Current$NestsNotFound
NC = Summary_SuccessByYearDataFrame_Current$NestsNotCounted
NewNests = Summary_SuccessByYearDataFrame_Current$New
LostNests = Summary_SuccessByYearDataFrame_Current$Lost
#YngPerOccupiedNest = signif(Summary_SuccessByYearDataFrame_Current$`Young/occupied nest`,3)
YngPerSuccessfulNest = signif(Summary_SuccessByYearDataFrame_Current$YoungPerSuccessfulNest,3)
PctOccupied = signif(Summary_SuccessByYearDataFrame_Current$PctOccupied,2)
PctSuccessful = signif(Summary_SuccessByYearDataFrame_Current$PctSuccessful,2)
ProductivityUnknown = Summary_SuccessByYearDataFrame_Current$`Productivity unknown`

ExecutiveSummaryText = paste("Nest occupancy surveys were flown between ",GetLongDate(OStart)," and ",GetLongDate(OEnd),", followed by productivity surveys between ",GetLongDate(PStart)," and ",GetLongDate(PEnd),". We observed ",NestsLocated," nests, of which ",Occupied," were occupied (",PctOccupied,"%). Productivity surveys revealed that ",Successful," occupied nests produced young (",PctSuccessful,"%). Productivity could not be determined for ",ProductivityUnknown," nests. Successful nests resulted in ",YoungFledged," young, an average of ",YngPerSuccessfulNest," young per successful nest. ", str_to_sentence(xfun::numbers_to_words(LostNests))," nests were lost and ", NewNests," new ",ifelse(NewNests == 1,"nest was","nests were")," discovered.",sep="")


```

# Survey Summary: Bald Eagle Monitoring in Wrangell-St. Elias National Park and Preserve, `r SurveyYear`

Kyle Cutting^1^ and Scott Miller^2^

^1^National Park Service, Wrangell St. Elias National Park and Preserve\
8 Richardson Hwy Mile 106\
Copper Center, AK 99573

^2^U.S. Department of the Interior\
Enterprise Data Division\
240 W. 5th Ave Anchorage\
Alaska 99501

`r CurrentDate`

# Abstract

The Copper River in Wrangell-St. Elias National Park and Preserve (WRST), Alaska, contains a high density of nesting Bald Eagles <i>(Haliaeetus leucocephalus)</i> and the National Park Service has identified Bald Eagles as a focal fauna species to monitor. Bald Eagles in WRST are a high profile species that are dependent upon many resources along the Copper River and nest at the northern edge of the their range. Bald Eagles are top-trophic level predators and they often respond quickly to changes in their environment by changing their breeding activities.

In `r SurveyYear` we conducted an aerial bald eagle productivity survey in and around the park. `r ExecutiveSummaryText`

**Keywords:** Bald Eagle, Monitoring, Population, Wrangell-St. Elias National Park and Preserve, Alaska, Abundance, Productivity, Occupancy *Haliaeetus leucocephalus*, Copper River

# Objectives

The objectives of the monitoring efforts are to:

-   Document the annual changes in occupancy of bald eagles,

-   Document the annual changes in productivity of bald eagle nests, and

-   Document any changes in distribution of bald eagle nests.

# Methods

The `r SurveyYear` bald eagle survey followed the same methods (Putera and Miller, 2018) that have been used since `r MinYear`. Five survey areas were defined along the Copper River and its major tributaries (Figure 1):

1.) The Upper Copper River (UCR, 80km) extends from Copper and Tanada lakes to the confluence of the Chistochina River. White spruce (*Picea glauca*) is the predominant tree species, with balsam poplar (*Populus balsamifera*) common on the lower stretches.

2.) The Middle Copper River (MCR, 179km) extends from the Chistochina River south to the confluence of the Chitina River.The MCR contains a mixture of both white spruce and balsam poplar, with poplar being slightly more common. The Copper River begins to braid along the MCR.

3.) The Lower Copper River (LCR) section (188km) extends from the Chitina River south to Miles Lake. The LCR is dominated by cottonwood along the river. The vegetation also begins to exhibit coastal influences, adding several coastal species including black cottonwood (*Populus trichocarpa*), Sitka spruce (*Picea sitchensis*), and mountain hemlock (*Tsuga mertensiana*). The Copper River is wide and more braided than the upper sections, with steep canyons on either side. Nesting begins consistently later along the LCR and is probably influenced by the numerous glaciers in the area which delay snow melt and maintain the area in a winter-like state several weeks after green up has begun to occur to the north.

4.) The Bremner River (45km) is surveyed from its confluence with the Copper River to the forks by Threemile Canyon, and has similar characteristics as the LCR.

5.) The Chitina River (116km) is surveyed from the confluence of the Copper River to where the river becomes braided above the Tana River. White spruce dominate the trees, with cottonwood becoming scarce in the upper reaches.

```{r echo=FALSE}

# Load shapefiles to be used in the maps later.
Parks = read_sf(paste(ShpPath,"GCS/AKParks_reprojected.shp",sep=""))
Rivers = read_sf(paste(ShpPath,"Alaska_Rivers_1M.shp",sep=""))
AK  = read_sf(paste(ShpPath,"GCS/AK_GCS_poly.shp",sep=""))

# Build a map of the survey area showing park boundaries, copper river and tributaries, etc.
ggplot() +
  
  # Blue rectangle to make the background of the map blue
  geom_rect(aes(xmin = -155,xmax = -140,ymin = 55,ymax = 70),fill = "steelblue3") +
  
  # Alaska outline
  geom_sf(data=AK,mapping=aes(), color='darkgray',fill='snow2') +
  
  # Park boundary
  geom_sf(data=Parks,mapping=aes(), color='yellow4',fill='yellow4',alpha=0.3) +
  
  # Rivers
  geom_sf(data=Rivers,mapping=aes(),color='steelblue3',fill='steelblue3') +
  
  # Set the map limits, zoom into the BAEA monitoring area
  coord_sf(xlim = c(-146.6, -142), ylim = c(60.5, 62.8), expand = FALSE) +
  xlab('Longitude') +
  ylab('Latitude')+
  
  # Minimal theme
  theme_minimal() +
  
  # WRST text Label
  geom_text(aes(x = -144.25, y = 61.8, label = "Wrangell - St. Elias National\nPark and Preserve"), size = 4, color = "gray40", fontface = "bold") +
  
  # Upper Copper River label
  geom_text(aes(x = -144, y = 62.45, label = "Upper Copper River"), size = 4, color = "black") +
  
  # Middle Copper River label
  geom_text(aes(x = -145.2, y = 61.5, label = "Middle Copper River"), size = 4, color = "black") +
  
  # Lower Copper River label
  geom_text(aes(x = -145.2, y = 60.8, label = "Lower Copper River"), size = 4, color = "black") +
  
  # Chitina River label
  geom_text(aes(x = -143.8, y = 61.25, label = "Chitina River"), size = 4, color = "black") +
  
  # Bremner River label
  geom_text(aes(x = -143.8, y = 60.95, label = "Bremner River"), size = 4, color = "black")

```

Figure `r FigureCounter`. Map of the bald eagles monitoring survey area in the Copper River drainage, Wrangell-St. Elias National Park and Preserve. Major tributaries in the survey area include the Chitina and Bremner rivers.

Two aerial surveys were conducted in `r SurveyYear` - one to document territorial occupancy, and one to document productivity. Surveys were flown using a small airplane flying at low level (\<100m AGL) and low speed (\<100 kph), with one experienced observer. Flight paths followed the river's edge, with the opposite side covered on the return trip. In most cases, the relatively narrow riparian corridor allowed adequate coverage with a single pass on each side of the river. Where the corridor widened, multiple passes were made to adequately evaluate the available habitat.

For all new nests located, the following information was collected:

a)  Location (latitude/longitude via handheld Global Positioning System receiver),
b)  Count of adult birds,
c)  Behavior of adults (incubating, perched),
d)  Number of eggs observed,
e)  Number of nestlings,
f)  Species of nest tree (Populus spp., Picea spp.),
g)  Whether the nest tree was alive or dead, and
h)  Nest status (good, poor, remnant, gone).

All information except latitude/longitude was recorded on field forms. All nests still intact from the previous year's surveys were checked for occupancy during the occupancy survey. Information recorded for located nests include:

a)  Count of adult birds,
b)  Behavior of adults, and
c)  Number of eggs or nestlings observed.

During the second survey, all nest locations collected during prior surveys were checked for productivity. Territorial occupancy was defined based on a bird in the nest in incubating posture or two adult birds at a nest or in the nest tree (Postupalsky, 1974). For the productivity survey, all nests observed with an incubating bird or two adult birds at a nest during the occupancy survey were checked to determine nest outcome. Multiple passes were commonly needed to count chicks, particularly in the Lower Copper area where cottonwoods are more common, and leaves make counting chicks difficult.

# Results

Bald eagle occupancy survey flights were conducted in WRST from `r GetLongDate(OStart)` to `r GetLongDate(OEnd)`. Follow-up productivity flights happened from `r GetLongDate(PStart)` to `r GetLongDate(PEnd)`. `r ExecutiveSummaryText`Table `r TableCounter` shows nest success and productivity metrics by river segment. Spatial distribution of surveyed nests and their productivity is shown in Figure `r FigureCounter`.

Table `r TableCounter`. Results summary, `r SurveyYear` bald eagle survey.

```{r echo=FALSE, label="Current survey metrics table"}
# Summarize the eagle counts for the current survey into 
# a table, transpose it so it fits the report better, modify the header, and output
Summary_SuccessByYearDataFrame_Current = Summary_SuccessByYearDataFrame %>% filter(Year == SurveyYear) %>% select(-Year,-Records)
Summary_SuccessByYearDataFrame_Current_char = sapply(Summary_SuccessByYearDataFrame_Current, as.character) 

# Convert it into a data frame
Summary_SuccessByYearDataFrame_Current_char_DF = as.data.frame(Summary_SuccessByYearDataFrame_Current_char)
colnames(Summary_SuccessByYearDataFrame_Current_char_DF) = 'Result'

#Output
kable(Summary_SuccessByYearDataFrame_Current_char_DF,col.names=c("Metric","Result"))

# Increment the table counter
TableCounter = TableCounter + 1

```

```{r echo=FALSE}

# Get the spatial locations of successful and failed nests for a map
Sql = paste("SELECT   
Nest_ID
, RiverSegment
,Latitude AS Latitude
, Longitude AS Longitude
, NumYng
,CASE 
  WHEN StatusCode_P = 'IF' THEN 'Failed' 
  WHEN StatusCode_P = 'S' THEN 'Succeeded' 
  ELSE NULL 
END AS Status
, StatusCode_P
, Nest_O
FROM Dataset_Productivity
WHERE (Year = ",SurveyYear,") 
AND (Latitude IS NOT NULL) -- Avoid missing spatial coords
AND (Latitude > 50) -- Avoid Lat/Lon records of 0/0
AND (StatusCode_P IN ('S', 'IF')) -- Successful or faild, ignore not found and not counted records
",sep='')
#cat(Sql)

# Get data frame from query
Dataset_ProductivityDataFrame = GetDataFrameFromDatabase(Sql)
Dataset_ProductivityDataFrame[is.na(Dataset_ProductivityDataFrame)] = " "

# Map Title
title = paste("Dispersion of successful nests, ", SurveyYear)

# Query out just the successful nests
#SuccessFulNestsDataFrame = subset(Dataset_ProductivityDataFrame,StatusCode_P == 'S')

# Plot the nest success on a map
ggplot() +
  
  # Blue rectangle to make the background of the map blue
  geom_rect(aes(xmin = -155,xmax = -140,ymin = 55,ymax = 70),fill = "steelblue3") +
  
  # Alaska outline
  geom_sf(data=AK,mapping=aes(), color='darkgray',fill='snow2') +
  
  # Park boundary
  geom_sf(data=Parks,mapping=aes(), color='yellow4',fill='yellow4',alpha=0.3) +

    # Rivers
  geom_sf(data=Rivers,mapping=aes(),color='steelblue3',fill='steelblue3') +  
  
  # Failed nests
  # Consider carefully whether to publish nest locations, or not
  geom_point(data=Dataset_ProductivityDataFrame, aes(x = Longitude, y = Latitude,colour=Status,shape=Status),size=2) +
  scale_color_manual(values = c("Succeeded" = "black", "Failed" = "red")) +
  scale_shape_manual(values = c(4,19)) + # 4 = X, 19 = dot
  
  # Productive nests, increasing size with productivity
  geom_point(data=subset(Dataset_ProductivityDataFrame,Status=="Succeeded"), aes(x = Longitude, y = Latitude, size=factor(NumYng)),color="black") +
  # Override the size attributes assigned by geom_point. 'values' is size of symbol.
  scale_size_manual(name="Young produced",values = c(2,5),breaks = c(1,2), labels = c("1","2")) + 
  
  # Set the map limits
  coord_sf(xlim = c(-146.6, -142), ylim = c(60.5, 62.8), expand = FALSE) +
  xlab('Longitude') +
  ylab('Latitude')+
  
  # Minimal theme
  theme_minimal() +
  
  # WRST Label
  geom_text(aes(x = -143.5, y = 61.8, label = "Wrangell - St. Elias 
National Park and Preserve"), size = 3, color = "gray40") +
  
    # Upper Copper River label
  geom_text(aes(x = -144, y = 62.45, label = "Upper Copper\nRiver"), size = 3, color = "dodgerblue4") +
  
  # Middle Copper River label
  geom_text(aes(x = -145.2, y = 61.5, label = "Middle Copper River"), size = 3, color = "dodgerblue4") +
  
  # Lower Copper River label
  geom_text(aes(x = -145.2, y = 60.8, label = "Lower Copper River"), size = 3, color = "dodgerblue4") +
  
  # Chitina River label
  geom_text(aes(x = -143.8, y = 61.25, label = "Chitina River"), size = 3, color = "dodgerblue4") +
  
  # Bremner River label
  geom_text(aes(x = -143.8, y = 60.95, label = "Bremner River"), size = 3, color = "dodgerblue4")

# Increment the figure counter
FigureCounter = FigureCounter + 1
```

Figure `r FigureCounter`. Bald eagle nest dispersion and breeding pair productivity metrics in the Copper River Valley in `r SurveyYear`. Black circles represent successful nests with the size indicating number of young produced. Failed nests are shown with an X. Mean occupancy rate was `r PctOccupied`% and mean success rate was `r PctSuccessful`%.

Bald eagle productivity metrics are presented by river segment in Table `r TableCounter`.

Table `r TableCounter`. Bald eagle productivity and success rates by river segment, `r SurveyYear`.

```{r echo=FALSE,label="Bald eagle productivity and success rates by river segment"}

# Retrieve the annual summary information from the database
Sql = paste("SELECT  * FROM DashboardByRiverSegment
WHERE Year <= ",SurveyYear," 
And RiverSegment <> 'Chitina River Tributaries' -- only done once
And NOT(RiverSegment = 'Chitina' And Year = 1989) -- Only two nests observed, one successful, inflates chitina productivity so omit
ORDER BY Year",sep="")

# Get the data from the database into a data frame
SuccessByYearRiverSegmentDF = GetDataFrameFromDatabase(Sql) 

# Filter the above data frame on the current year and omit the first (Year) column
CurrentSuccessByYearRiverSegmentDF = SuccessByYearRiverSegmentDF %>%
  filter(Year == SurveyYear) %>%
  select(RiverSegment,
         AvailableNests,
         AvailableButOmitted,
         `Nests located`,
         `Nests not located`,
         `Productivity unknown`,
         `Productivity known`,
         New,Lost,NetGain,
         Occupied,Successful,
         PctOccupied,PctSuccessful,
         Young,YoungPerSuccessfulNest
)

# Output the data
kable(t(CurrentSuccessByYearRiverSegmentDF))

# Increment the table counter
TableCounter = TableCounter + 1
```

```{r,echo=FALSE}
df_long <- CurrentSuccessByYearRiverSegmentDF %>% 
  select(RiverSegment,PctOccupied,PctSuccessful) %>% 
  #mutate(Failed = 100 - (PctOccupied + PctSuccessful)) %>%
  pivot_longer(
    cols = c(PctOccupied, PctSuccessful),
    names_to = "Category",
    values_to = "Value"
  ) %>%
  mutate(
    Category = recode(Category,
                      PctOccupied = "Percent Occupied",
                      PctSuccessful = "Percent Successful"))

# Stacked + percent
ggplot(df_long, aes(fill = Category, y = Value, x = RiverSegment)) +
  geom_bar(position = "dodge", stat = "identity") +
  theme_minimal() +
  theme(
    legend.position = "top")


```

Figure `r FigureCounter`. Relative percent nest occupancy and succes by river segment in the Copper River valley, `r SurveyYear`.

# Reference

<!-- Kozie K. 1996. [Bald Eagle Productivity for Wrangell-St Elias National Park and Preserve 1993-1996](https://irma.nps.gov/DataStore/Reference/Profile/2238328). National Park Service, Wrangell - St. Elias National Park. Unpublished Report. Copper Center, Alaska -->

Postupalsky, S. 1974. Raptor reproductive success: some problems with methods, criteria, and terminology, IN Raptor Research Report No. 2, Raptor Research Foundation, Vermilion, South Dakota, pp 21-31.

Putera, J. and S.D. Miller. 2018. [Protocol implementation plan for bald eagle monitoring in Wrangell-St. Elias National Park & Preserve, Central Alaska Network: Version 1.0](https://irma.nps.gov/DataStore/Reference/Profile/2256539). Natural Resource Report. NPS/CAKN/NRR—2018/1752. National Park Service. Fort Collins, Colorado.

```{r echo=FALSE,include=FALSE}
# This chunk copies the report to the J or O drive
CurrentDirectory = "C:/Development/WRST BAEA Reports R Markdown/"
From = paste(CurrentDirectory,"WRST-BAEA-Annual-Summary-Report.html",sep='')
To = paste(CurrentDirectory,SurveyYear," WRST BAEA Survey Summary.html",sep='') # ARCN

# Copy the file to the Network reports directory
file.copy(From, To, overwrite = TRUE)

```

<!-- ## R Markdown Report Template -->

<!-- This template was generated automatically from an R Markdown script available at <https://github.com/NPS-ARCN-CAKN/WRST-BAEA-Annual-Report-R-Markdown>. The information in the report was generated from bald eagle survey data in the NPS WRST_Caribou SQL Server database (inpyugamsvm01\nuna:WRST_Caribou). -->
